---
title: "Stat 6021: Module B Practice Questions And Answers"
author: "Tom Lever"
date: 08/20/22
output:
  pdf_document: default
  html_document: default
---


# Topic B.2: Getting Started with R

1. The `mtcars` dataset comes built-in with R. The dataset was extracted from the 1974 *Motor Trend* US magazine and comprises several characteristics of automobile design and performance for 32 automobiles (1973 to 1974 models).

    a) Read this data in and assign it to an object called `cars.df`.
    
        ```{r, echo=TRUE, results="show"}
        cars.df <- mtcars
        ```
        
    b) Use the environment window and report how many observations and variables are in this dataset.
    
        ```{r, echo=TRUE, results="show"}
        number_of_observations <- nrow(cars.df)
        number_of_observations
        number_of_variables <- ncol(cars.df)
        number_of_variables
        ```

2. Download the text tile `students.txt` from Collab. The file's dataset contains information on students taking an introductory statistics class at a large public university.

    a) Read this data in and assign it to an object called `students.df`.
    
        ```{r, echo=TRUE, results="show"}
        students.df <- read.table("students.txt", header=TRUE)
        ```
        
    b) Use the environment window and report how many observations and variables are in this dataset.
    
        ```{r, echo=TRUE, results="show"}
        number_of_observations <- nrow(students.df)
        number_of_observations
        number_of_variables <- ncol(students.df)
        number_of_variables
        ```
        
3. Install the following packages:

    a) tidyverse
    
        ```{r, echo=TRUE, results="show"}
        install.packages("tidyverse", repos = "http://cran.us.r-project.org")
        ```
        
    b) faraway
    
        ```{r, echo=TRUE, results="hide"}
        install.packages("faraway", repos = "http://cran.us.r-project.org")
        ```
    
    c) MASS
    
        ```{r, echo=TRUE, results="hide"}
        install.packages("MASS", repos = "http://cran.us.r-project.org")
        ```
    
    d) leaps
    
        ```{r, echo=TRUE, results="hide"}
        install.packages("leaps", repos = "http://cran.us.r-project.org")
        ```
    
    e) ROCR
    
        ```{r, echo=TRUE, results="hide"}
        install.packages("ROCR", repos = "http://cran.us.r-project.org")
        ```
    
    f) nycflights13
    
        ```{r, echo=TRUE, results="hide"}
        install.packages("nycflights13", repos = "http://cran.us.r-project.org")
        ```
    
    g) gapminder
    
        ```{r, echo=TRUE, results="hide"}
        install.packages("gapminder", repos = "http://cran.us.r-project.org")
        ```
    
    h) palmerpenguins
    
        ```{r, echo=TRUE, results="hide"}
        install.packages("palmerpenguins", repos = "http://cran.us.r-project.org")
        ```
    
4. Load the `faraway` package, and read in the dataset called cornnit and assign it to an object called `corn.df`.

    ```{r, echo=TRUE, results="show"}
    library("faraway")
    corn.df <- cornnit
    ```

# Topic B.1: Getting Started with R

5. Are the following valid names for objects in R?

    a) 2020_Major
    
        No
        
    b) .2020.Age
    
        No
       
    c) #Courses.2020
    
        No
        
    d) _courses_2020
    
        No
        
    e) Fav_Sport20
    
        Yes
        
    f) major 2020
    
        No
    
    g) age(2020)
    
        No
        
    h) FavSport_2020
    
        Yes
        
6. Create a numeric vector with the following 10 values.

    \begin{center}
        13, 91, 36, 95, 9, 3, 61, 20, 22, 97
    \end{center}
    
    Assign these ten values to a vector called `practice`. Use R to find out if `practice` is a character, numeric, or logical object.
    ```{r, echo=TRUE, results="show"}
    practice <- c(13, 91, 36, 95, 9, 3, 61, 20, 22, 97)
    class(practice)
    ```
    
7. For each of the following, determine if they are `TRUE` or `FALSE`. Try answering first without using R, then use R to confirm.

    a) practice[5] == 5
    
        ```{r, echo=TRUE, results="show"}
        practice[5] == 5
        ```
    
    b) practice[10] != 97
    
        ```{r, echo=TRUE, results="show"}
        practice[10] != 97
        ```
    
    c) (practice[1] + practice[2]) < 104
    
        ```{r, echo=TRUE, results="show"}
        (practice[1] + practice[2]) < 104
        ```
    
    d) (practice[1] + practice[2]) <= 104
    
        ```{r, echo=TRUE, results="show"}
        (practice[1] + practice[2]) <= 104
        ```
    
    e) (practice[2] == 91) & (practice[9] == 22)
    
        ```{r, echo=TRUE, results="show"}
        (practice[2] == 91) & (practice[9] == 22)
        ```
    
    f) (practice[5] < 9) | (practice[6] >= 4)
    
        ```{r, echo=TRUE, results="show"}
        (practice[5] < 9) | (practice[6] >= 4)
        ```

8. Create the following matrix in R, and assign it to the object Mat.A.

    \begin{center}
        \[
            \begin{bmatrix}
                4 & 1 & 3 \\
                6 & 2 & 1
	           \end{bmatrix}
	       \]
    \end{center}
    
    ```{r, echo=TRUE, results="show"}
    Mat.A <- cbind(c(4,6), c(1,2), c(3,1))
    ```
    
    a) Give the following column names to Mat.A: "Huey, "Dewey", "Louie".
        
        ```{r, echo=TRUE, results="show"}
        colnames(Mat.A) <- c("Huey", "Dewey", "Louie")
        Mat.A
        ```
    
    b) Without using R, what is the output if we type Mat.A[2,1]?
    
        ```{r, echo=TRUE, results="show"}
        Mat.A[2,1]
        ```
     
    c) Without using R, what is the output if we type dim(Mat.A)?
    
        ```{r, echo=TRUE, results="show"}
        dim(Mat.A)
        ```
        
9. Convert the vector practice to a factor. What is the order of the levels in this factor?
        
    ```{r, echo=TRUE, results="show"}
    factor(practice)
    length(factor(practice))
    ```
    
# Topic B.4: R Markdown

10. Type up your answers to these questions using R Markdown, and output an HTML file.

    ```{r, eval=FALSE, echo=TRUE, results="show"}
    rmarkdown::render("Practice_Answers.Rmd", "html_document")
    ```