---
title: "Stat 6021: Addressing Guided Question Set 2"
author: "Tom Lever"
date: 09/02/22
output:
  pdf_document: default
urlcolor: blue
linkcolor: red
---

For this exercise, use the `.csv` data file that you created at the end of the previous guided question set, `new_students.csv`.
As a reminder, the dataset contains information on students taking an introductory statistics class at a large public university.
The columns of data are:

    * `Student`: ID number on survey
    * `Gender`: gender of student (male / female)
    * `Smoke`: whether the student smokes (yes / no)
    * `Marijuan`: whether the student smokes marijuana (yes / no)
    * `DrivDrnk`: whether the student has ever driven while drunk (yes / no)
    * `GPA`: student's current GPA
    * `PartyNum`: number of days per month the student parties
    * `DaysBeer`: number of days per month the student has at least two alcoholic drinks
    * `StudyHrs`: number of hours spent studying per week
    * `PartyAnimal`: whether the students parties more than 8 days per month (yes / no)
    * `GPA.cat`: "low" if GPA is less than 3.0, "moderate" if GPA is at least 3.0 and less than 3.5, "high" if GPA is at least 3.5

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
students_dataframe <- read.csv(
    file = "../../Module_1--Data_Wrangling/Guided_Question_Set/new_students.csv"
)
head(students_dataframe, n = 3)
```

1. Produce a frequency table of the number of students in each level of `GPA.cat`.
If needed, be sure to arrange the order of the output appropriately.
How many students are in each level of `GPA.cat`?

    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    table(factor(students_dataframe$GPA.cat, labels = c("low", "moderate", "high")))
    ```