---
title: "Stat 6021: Project 1"
author: "Group 2: Ben G. Ballard, Sirish Kumar Desai, Kevin Kuc, Tom Lever"
date: 10/01/22
output:
  pdf_document: default
  html_document: default
urlcolor: blue
linkcolor: red
---

# Executive Summary

Blue Nile, Inc. contacted our team of data analysts to review a data set containing data relating to $1214$ various diamonds for sale on Blue Nile's website. Our team was tasked with exploring the relationships between key attributes of the diamonds: weight (carat); clarity; color; cut; and price. Below, we further describe these data and variables and present visualizations and analysis. Our visualizations illustrate proportions of diamonds across groups of variables. Our analysis describes trends within the visualizations and evaluates claims on Blue Nile's website.

Cut refers to how ideally a diamond is shaped, and consists of a range from $Good$ to $Astor \ Ideal$. $Astor \ Ideal$ diamonds have the highest median price followed by $Ideal$, $Very \ Good$, and $Good$ diamonds. Blue Nile implies that $Astor \ Ideal$ diamonds gather and reflect the most light possible, and may on this basis give $Astor \ Ideal$ diamonds relatively high prices.

Color is the second most significant factor that influences the price of a diamond. As the diamond becomes more colorful the diamond generally decreases in value. The Gemological Institute of America (GIA) color scale is the industry standard for diamond grading. The scale starts with $D$ for least colorful diamonds and alphabetically proceeds to $Z$ for most colorful diamonds (although Blue Nile only carries diamonds with color identifiers in the range $D$ through $J$).

The third factor that affects a diamonds price is clarity, which is an assessment of small imperfections on the surface and within the stone. The clarity of a diamond ranges from $SI2$, which has inclusions observable by a human under $10x$ magnification, to an $FL$ diamond, which is flawless. Flawless diamonds have the highest median price among diamonds grouped by clarity. Their lack of abundance drives up their price; less than $1$ percent of all diamonds have $FL$ clarity. The $VS1$ clarity identifier corresponds to a group of diamonds with the second highest median price, although four groups of diamonds have greater clarity.

The weight of a diamond is measured in carats; Blue Nile and others refer to the weight of a diamond as carat weight and carat. The weight of a diamond is not always indicative of a diamond's size. That being said, we found that as weight increases, the price of a diamond also increases, at an increasing rate, with increasing variability. In "Simple Linear Regression of Price versus Weight", we determine a linear model that relates transformed price to transformed weight. This linear model may be used to predict diamond price given weight.
For an increase in weight of $10$ percent, price increases by a factor of $1.2$. For example, for an increase in weight
from $1 \ carat$ to $1.1 \ carat$, price increases from $\$5,019.07$ by a factor of $1.2$ to $\$6,040.75$. These prices have a ratio of $1.2$.

Overall, the cut, color, clarity, and carat of a diamond are all predictors for a diamond's market value,
with various qualifications.

# Presentation and Analysis of Data

## Description of Data and Variables

Our data set describes $1214$ difference diamonds that are for sale at [http://www.bluenile.com](http://www.bluenile.com).
Our data set describes diamonds with carat, clarity, color, cut, and price data.
Our data set describes a subset of the diamonds that are for sale with a subset of features.
Table 1 presents data for three diamonds.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE}
library(dplyr)
library(kableExtra)
data_set <- read.csv("diamonds4.csv", header = TRUE) %>% rename(weight = carat) %>% mutate(clarity = factor(clarity, c("SI2", "SI1", "VS2", "VS1", "VVS2", "VVS1", "IF", "FL")), color = factor(color, c("J", "I", "H", "G", "F", "E", "D")), cut = factor(cut, c("Good", "Very Good", "Ideal", "Astor Ideal")))
first_three_rows <- head(data_set, n = 3)
kbl(first_three_rows, caption = "first three diamonds in our data set", booktabs = TRUE, longtable = TRUE) %>% kable_styling(latex_options = "striped", position = "center")
```

Weight measures a diamond's weight in carats.
Clarity assesses small imperfections within a diamond and quantifies and specifies inclusions.
Color refers to how colorless a diamond is.
Cut measures how well-proportioned a diamond's dimensions are.

## General Analysis and Addressing Claims on BlueNile's Website

### General Analysis

This report explores the relationships between key attributes of diamonds: specifically,
weight, cut, color, clarity, and price.

We analyze the data one variable at a time through univariate data visualizations,
including bar charts, box plots, histograms, and plots of probability-density distributions.
These visualizations depict proportions of diamonds versus values of individual variables,
summaries of values of individual variables, and discrete and continuous probability density distributions
versus values of individual variables. Proportions, descriptive statistics, probabilities, and
information about range, spread, and skewness may be determined from this visualizations.

We use bivariate visualizations to quickly understand the relationships between pairs of variables.
For example, we considered the relationships of price and cut, price and color, price and clarity, and 
price and weight. Price was clearly higher for some values of cut, color, clarity, and weight,
or clearly followed trends across values of cut, color, clarity, or weight.
Flawlessly clear diamonds had a much higher median price than diamonds of any other clarity.
Sometimes, the relationship between price and cut, color, clarity, or weight was let obvious.
Diamonds of one particular color do not stand out as having a higher price than diamonds of another color.

We use multivariate visualizations to understand the relationships between a variable and
multiple other variables, as well as to compare the relationships between a variable and other variables.
For example, we considered the relationships of price and various combinations of other variables.

Our bivariate and multivariate visualizations include boxplots and scatterplots.

Over the course of our analysis of our visualizations, we evaluated claims on BlueNile's website,
which are presently immediately below.

Our team developed a simple linear regression model for price versus weight to deeply understand
the relationship between price and weight and to provide a means of predicting price given weight.
In developing this model, we identified five unmet assumptions of simple linear regression.
We met these assumptions through further data analysis and data transformation.
Our modeling method may be extended to understanding the relationship between any two quantitative
variables.

### Addressing Claims on BlueNile's Website

Considering the claim at [https://www.bluenile.com/education/diamonds](https://www.bluenile.com/education/diamonds)
that "The higher [in idealness] a diamond is, the more expensive it will be",
this claim is supported by a trend that as diamonds improve in cut, third-quartile price increases;
a trend that as diamonds improve in cut from good to very good to ideal, maximum price increases;
and the fact that Astor ideal diamonds have the highest minimum / first-quartile / median / third-quartile price.
However, this claim is refuted by a trend that as diamonds improve in cut from good to very good to ideal,
minimum / first-quartile / median price decreases.

Considering the claim that "The higher [in clarity] a diamond is, the more expensive it will be",
this claim is supported in that flawless diamonds have the highest first quartile / median / third-quartile / maximum price.
However, there is no trend across other clarities of diamond. There may even be a slight negative trend considering medians and third quartiles.

Considering the claim that "The higher [in color] a diamond is, the more expensive it will be",
this claim is supported by a trend that for a transition from a group of diamonds with a color identifier in the set
{"G", "F", "E", "D"} to a group of diamonds with a color identifier closer to "D",
the min / first-quartile / median / third-quartile / max price increases.

Considering the claim that "The higher [in weight] a diamond is, the more expensive it will be,"
we find in a scatterplot of price versus weight a trend that s weight of a diamond increases,
price of the diamond increases, at an increasing rate.

Considering a claim that cut is the most significant and a claim that color is second-most significant,
color may actually be more significant than cut, given graphs of price versus weight versus color and price versus weight versus cut.
For the same scales of price and weight, most changes between colors result in significant changes in slope,
whereas most changes between cuts result in less significant changes in slope.
Additionally, considering high prices and weights, the most colorless diamonds have the highest prices and weights,
whereas Astor ideal diamonds have the lowest prices and weights.

## Detailed Visualizations and Plot-Specific Analysis

Considering proportions of diamonds by clarity, the group of diamonds with clarity identifier $SI1$
has the highest proportion of about $0.2$. For a transition from a group of diamonds with a clarity identifier in the set
{$VS1$, $VVS2$, $VVS1$, $IF$, $FL$} to a group of diamonds with clarity identifier closer to $FL$,
proportion of diamonds decreases. The group of diamonds with clarity identifier $FL$ has the lowest proportion of less than $0.01$.

Considering proportions of diamonds by color, the group of diamonds with color identifier $F$
has the highest proportion of over $0.175$. The group of diamonds with color identifier $J$
has the lowest proportion of about $0.075$. There seems to be a general trend of proportion increasing
as color identifier approaches $D$.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, figures-side, fig.width=3.3, fig.height=2.4}
library(ggplot2)
clarity_and_proportion <- data_set %>% select(clarity) %>% group_by(clarity) %>% summarize(count = n()) %>% mutate(proportion = count / nrow(data_set))
ggplot(clarity_and_proportion, aes(x = clarity, y = proportion)) +
    geom_bar(stat = "identity", fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Diamonds by Clarity") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 90)
    )

color_and_proportion <- data_set %>% select(color) %>% group_by(color) %>% summarize(count = n()) %>% mutate(proportion = count / nrow(data_set))
ggplot(color_and_proportion, aes(x = color, y = proportion)) +
    geom_bar(stat = "identity", fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Diamonds by Color") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering proportions of diamonds by cut, the group of ideal diamonds has the highest proportion of over $0.6$.
The group of Astor ideal diamonds has the lowest proportion of well under $0.005$.
As cut improves from good to ideal, proportion of diamonds increases.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4, fig.align='center'}
library(ggplot2)
cut_and_proportion <- data_set %>% select(cut) %>% group_by(cut) %>% summarize(count = n()) %>% mutate(proportion = count / nrow(data_set))
ggplot(cut_and_proportion, aes(x = cut, y = proportion)) +
    geom_bar(stat = "identity", fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Diamonds by Cut") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering a boxplot and a probability density distribution for weight,
the middle $50$ percent of weights fall between about $0.4 \ carat$ and $1.0 \ carat$.
The distribution is non-normal and positively / right skewed.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(y = weight)) +
    geom_boxplot(fill = "Blue", color = "Orange") +
    labs(
        y = "weight",
        title = "Boxplot of Weight"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank()
    )

ggplot() +
    geom_histogram(data = data_set, aes(x = data_set$weight, y = ..density..), binwidth = density(data_set$weight)$bw, fill = "Blue", color = "Orange") +
    geom_density(data = data_set, aes(x = data_set$weight), color = "Orange") +
    labs(
        x = "weight",
        y = "probability density",
        title = "Histogram and Probability Density\nDistribution of Weight"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering a boxplot and a probability density distribution for price,
the middle $50$ percent of logarithms of prices fall between 6.5 and 7.5.
The distribution is non-normal and positively / right skewed.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
library(ggplot2)
ggplot(data_set, aes(y = log(price))) +
    geom_boxplot(fill = "Blue", color = "Orange") +
    labs(
        y = "logarithmicized price",
        title = "Boxplot of Logarithmicized Price"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank()
    )

ggplot() +
    geom_histogram(data = data_set, aes(x = log(data_set$price), y = ..density..), binwidth = density(log(data_set$price))$bw, fill = "Blue", color = "Orange") +
    geom_density(data = data_set, aes(x = log(data_set$price)), color = "Orange") +
    labs(
        x = "logarithmicized price",
        y = "probability density",
        title = "Histogram and Probability Density\nDistribution of Logarithmicized Price"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering the relationship of price of a diamond in our data set and the cut of the diamond,
we construct boxplots of price versus cut with and without outliers and present the boxplot without outliers.
Including outliers, a very good diamond has the highest price of over $\$350,000$.
Excluding outliers, an ideal diamond has the highest price at about $\$10,500$.
A very good diamond has the lowest price at $\$322$.
The minimum / first-quartile / median price and interquartile range of prices of Astor ideal diamonds are highest.
The minimum / first-quartile / median prices of good, very good, and ideal diamonds decrease in that order,
and are less than the minimum / first-quartile / median price of Astor ideal diamonds.
The third-quartile price of Astor ideal diamonds is highest at about $\$6,500$.
The third-quartile prices of good, very good, and ideal diamonds increase in that order,
and are less than the third-quartile price of Astor ideal diamonds.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4, fig.align='center'}
ggplot(data_set, aes(x = cut, y = price)) +
    geom_boxplot(outlier.shape = NA, fill = "Blue", color = "Orange") +
    coord_cartesian(ylim = c(0, 1.125*10^4)) +
    labs(title = "Distributions of Price by Cut") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
minimum_price <- min(data_set$price)
index_of_minimum_price <- match(minimum_price, data_set$price)
cut_corresponding_to_minimum_price <- data_set$cut[index_of_minimum_price]
#minimum_price
#cut_corresponding_to_minimum_price
```

Considering the relationship of price of a diamond in our data set and the color of the diamond,
we construct boxplots of price versus color with and without outliers and present the boxplot without outliers.
Including outliers, for a transition from a group of diamonds with a color identifier in the English alphabet
to a group of diamonds with a color identifier of the next letter closer to the beginning of the alphabet, maximum price of a diamond increases
from around $\$50,000$ to around $\$350,000$.
Excluding outliers, a group of diamonds with color identifier $H$ has the highest price of about $\$13,750$
and the highest interquartile range of prices.
A diamond with color identifier $D$ has the lowest price of $\$322$.
For a transition from a group of diamonds with a color identifier $G$ to a group of diamonds with a color identifier of a letter
closer to the beginning of the alphabet, the first-quartile / median / third-quartile price of a diamond increases.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(x = color, y = price)) +
    geom_boxplot(fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Price by Color") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )

ggplot(data_set, aes(x = color, y = price)) +
    geom_boxplot(outlier.shape = NA, fill = "Blue", color = "Orange") +
    coord_cartesian(ylim = c(0, 1.5*10^4)) +
    labs(title = "Distributions of Price by Color") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
color_corresponding_to_minimum_price <- data_set$color[index_of_minimum_price]
#color_corresponding_to_minimum_price
```

Considering the relationship of price of a diamond in our data set and the clarity of the diamond,
we construct boxplots of price versus clarity with outliers and group of diamonds with clarity identifier $FL$,
and without outliers and without the group of diamonds with clarity identifier $FL$.
Including outliers and the group of diamonds with clarity identifier $FL$,
a diamond with clarity identifier $FL$ has the highest first-quartile / median / third-quartile / maximum price and interquartile range
of prices.
Excluding outliers and the group of diamonds with clarity identifier $FL$,
a diamond with clarity identifier $VS1$ has the lowest price.
A diamond with clarity identifier $VS1$ has the highest first-quartile / third-quartile / maximum price and interquartile range of prices.
A diamond with clarity identifier $SI1$ has the highest median price.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(x = clarity, y = price)) +
    geom_boxplot(fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Price by Clarity") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 90)
    )

ggplot(data_set %>% filter(clarity != "FL"), aes(x = clarity, y = price)) +
    geom_boxplot(outlier.shape = NA, fill = "Blue", color = "Orange") +
    coord_cartesian(ylim = c(0, 1.5*10^4)) +
    labs(title = "Distributions of Price by Clarity") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 90)
    )
#clarity_corresponding_to_minimum_price <- data_set$clarity[index_of_minimum_price]
#clarity_corresponding_to_minimum_price
```

Considering the relationship of weight of a diamond in our data set and the cut of the diamond,
we construct a boxplot of weight versus cut.
Including outliers, an ideal diamond has the highest weight of over $5 \ carat$.
Excluding outliers, an ideal diamond the highest weight of about $2 \ carat$.
An Astor ideal diamond has the lowest weight of $0.23 \ carat$.
Astor ideal diamonds have the highest first-quartile, median, and third-quartile weights.
Ideal diamonds have the greatest interquartile range of weights.
The median weights of good, very good, and ideal diamonds decrease in that order.

Considering the relationship of weight of a diamond in our data set and the color of the diamond,
we construct a boxplot of weight versus color.
Including outliers, a diamond with color identifier $E$ has the highest weight of over $7 \ carat$.
Excluding outliers, a diamond with color identifier $F$ or $E$ has the highest weight of around $2 \ carat$.
A diamond with color identifier $F$ has the lowest weight of $0.23 \ carat$.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(x = cut, y = weight)) +
    geom_boxplot(fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Weight by Cut") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
minimum_weight <- min(data_set$weight)
index_of_minimum_weight <- match(minimum_weight, data_set$weight)
cut_corresponding_to_minimum_weight <- data_set$cut[index_of_minimum_weight]
#minimum_weight
#cut_corresponding_to_minimum_weight

ggplot(data_set, aes(x = color, y = weight)) +
    geom_boxplot(fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Weight by Color") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
#color_corresponding_to_minimum_weight <- data_set$color[index_of_minimum_weight]
#color_corresponding_to_minimum_weight
```

Considering the relationship of weight of a diamond in our data set and the clarity of the diamond,
we construct a boxplot of weight versus color.
Including outliers, a diamond with clarity identifier $VS2$ has the highest weight of over $7 \ carat$.
Excluding outliers, a diamond with clarity identifier $FL$ has the highest weight over about $5.3 \ carat$.
A diamond with clarity identifier $VS1$ has the lowest weight of $0.23 \ carat$.
Diamonds with clarity identifier $FL$ have the highest first-quartile / median / third-quartile weights.
For a transition from a clarity identifier in the set {$SI2$, $SI1$, $VS2$, $VS1$, $VVS2$, $VVS1$, $IF$} to a clarity identifier closer to $IF$,
Weight generally decrease as clarity increases.

Considering the relationship of cut of a diamond in our data set and the color of the diamond,
we construct a bar chart of proportion of cut by color.
For each group of diamonds with a unique color identifier, most diamonds were ideal.
The proportions of ideal, very good, good, and Astor ideal diamonds decreased in that order.
The group of diamonds with color identifier $F$ had the highest proportion of Astor ideal diamonds.
The group of diamonds with color identifier $I$ had the highest proportion of ideal diamonds.
The group of diamonds with color identifier $G$ had the highest proportion of very good diamonds.
The group of diamonds with color identifier $D$ had the highest proportion of good diamonds.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(x = clarity, y = weight)) +
    geom_boxplot(fill = "Blue", color = "Orange") +
    labs(title = "Distributions of Weight by Clarity") +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 90)
    )
clarity_corresponding_to_minimum_weight <- data_set$clarity[index_of_minimum_weight]
#clarity_corresponding_to_minimum_weight

ggplot(data_set, aes(x = color, fill = cut)) +
    geom_bar(position = "fill") +
 labs(
     y = "proportion of cut",
     title = "Proportion of Cut by Color"
 ) +
 theme(
     plot.title = element_text(hjust = 0.5, size = 11),
     axis.text.x = element_text(angle = 0)
 )
```

Considering the relationship of cut of a diamond in our data set and the clarity of the diamond,
we construct a bar chart of proportion of cut by clarity.
For a group of diamonds with a clarity identifier other than $FL$, most diamonds are ideal.
For a transition from a group of diamonds with a clarity identifier in the set of identifiers {$VS1$, $VVS2$, $VVS1$, $IF$}
to a group of diamonds with a clarity identifier in that set closer to $IF$, the proportion of ideal diamonds increases
and the proportion of very good diamonds decreases, and the proportion of Astor ideal diamonds decreases.
For a transition from a group of diamonds with a clarity identifier in the set of identifiers {$SI2$, $SI1$, $VS2$, $VS1$, $VVS1$}
to a group of diamonds with a clarity identifier in that set closer to $VVS1$, the proportion of good diamonds decreases.
The majority of diamonds with clarity identifier $FL$ are very good, while all other diamonds with clarity identifier $FL$ are ideal.

Considering the relationship of color of a diamond in our data set and the clarity of the diamond,
we construct a bar chart of proportion of color by clarity.
The proportion of diamonds with color identifier $D$ for the group of diamonds with clarity identifier $FL$
is greater than $0.5$, is the only proportion greater than $0.5$, and is significantly greater than the proportion of diamonds
with color identifier $D$ for any other group of diamonds by clarity identifier.
For the group of diamonds with clarity identifier $FL$, all diamonds other than diamonds with color identifier $D$ have color identifier $F$.
For a transition from a group of diamonds with a clarity identifier in the set of identifiers {$VVS2$, $VVS1$, $IF$, $FL$}
to a group of diamonds with a clarity identifier in that set closer to $FL$, the proportion of diamonds with color identifier $F$ increases.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(x = clarity, fill = cut)) +
    geom_bar(position = "fill") +
 labs(
     y = "proportion of cut",
     title = "Proportion of Cut by Clarity"
 ) +
 theme(
     plot.title = element_text(hjust = 0.5, size = 11),
     axis.text.x = element_text(angle = 90)
 )

ggplot(data_set, aes(x = clarity, fill = color)) +
    geom_bar(position = "fill") +
 labs(
     y = "proportion of color",
     title = "Proportion of Color by Clarity"
 ) +
 theme(
     plot.title = element_text(hjust = 0.5, size = 11),
     axis.text.x = element_text(angle = 90)
 )
```

Considering the relationship of price, weight, and cut, in general, Astor ideal diamonds have the highest price at low weights,
and the lowest price at high weights. Sensitivity to weight decreases as cut improves. It may be that making the "most ideal cut to allow
light reflecting", or ensuring brilliance, requires removing significant material, or small diamonds.

Considering the relationship of price, weight, and color, the price of the most colorless diamonds increases more rapidly than for other
diamonds. There seems to be a general trend that as colorlessness increases, the price of diamonds, and the sensitivity of price, increase.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
ggplot(data_set, aes(x = weight, y = price, color = cut)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE) +
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Cut"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering the relationship of price, weight, and clarity, the price of flawless diamonds increases more rapidly than for other diamonds,
though there are only three flawless diamonds in our data set, two of the three have low weights, and the third has a high weight and high
price. There seems to be a general trend that as clarity increases, the price of diamonds, and the sensitivity of price, increase.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(x = weight, y = price, color = color)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE) +
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Color"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )

ggplot(data_set, aes(x = weight, y = price, color = clarity)) +
    geom_point() +
    geom_smooth(method = lm, se = FALSE)+
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Clarity"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering the relationship of price, weight, cut, and color, in general, colorless ideal and Astor ideal diamonds are priced highest
and have the greatest price sensitivity. Colorful good and very diamonds are priced lowest and have the lowest price sensitivity.
As color and cut improve, price and price sensitivity increase.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
ggplot(data_set, aes(x = log(weight), y = log(price), color = color, shape = cut)) +
    geom_point() +
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Cut vs. Color"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering the relationship of price, weight, cut, and clarity, in general, flawless ideal and Astor ideal diamonds are priced highest
and have the greatest price sensitivity. Good and very good diamonds with inclusions are priced lowest and have the lowest price sensitivity.
As cut and clarity improve, price and price sensitivity increase.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
ggplot(data_set, aes(x = log(weight), y = log(price), color = clarity, shape = cut)) +
    geom_point() +
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Cut vs. Clarity"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering the relationship of price, weight, color, and clarity, in general, colorless flawless diamonds are priced highest
and have the greatest price sensitivity. Colorfull diamonds with inclusions are priced lowest and have the lowest price sensitivity.
As color and clarity improve, price and price sensitivity increase.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.align='center'}
ggplot(data_set, aes(x = log(weight), y = log(price), color = color, shape = clarity)) +
    geom_point() +
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Color vs. Clarity"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering the relationship of price, weight, cut, color, and clarity, as clarity increases, colorless ideal diamonds
become relatively high priced. As color increases, ideal diamonds with minimal inclusions become relatively high priced.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
ggplot(data_set, aes(x = log(weight), y = log(price), color = color, shape = cut)) +
    geom_point() +
    facet_wrap(vars(clarity)) +
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Cut vs. Color vs. Clarity"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )

ggplot(data_set, aes(x = log(weight), y = log(price), color = clarity, shape = cut)) +
    geom_point() +
    facet_wrap(vars(color)) +
    labs(
        x = "weight",
        y = "price",
        title = "Price vs. Weight vs. Cut vs. Clarity vs. Color"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )
```

Considering the relationship of price of a diamond in our data set and the weight of the diamond,
we construct a scatterplot for a linear model of price versus weight,
as well as a plot of externally studentized residuals versus predicted prices for the model,
a plot of AutoCorrelation Function values versus lag for the model, and
a plot of sample quantiles versus theoretical quantiles for the residuals of the model.

As weight of a diamond increases, price of the diamond increases, at an increasing rate.
As weight of a diamond increases, the variance of residuals of a linear model of price versus weight increases.
There is a significant correlation between $(weight, \ price)$ observations and observations three away.
Given a moderate downward and a sharp upward curve at extremes of a plot of sample quantiles versus theoretical quantiles
for the residuals of the linear model of price versus weight, the tails of a probability vs. externally studentized residuals plot / distribution are too light for this distribution to be considered normal.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
ggplot(data_set, aes(x = weight, y = price)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
        x = "weight (carat)",
        y = "Price ($)",
        title = "Price vs. Weight"
    ) +
    theme(
    plot.title = element_text(hjust = 0.5, size = 11),
    axis.text.x = element_text(angle = 0)
    )

linear_model_of_price_and_weight <- lm(price ~ weight, data = data_set)

ggplot(
    data.frame(
        externally_studentized_residual = linear_model_of_price_and_weight$residuals,
        predicted_price = linear_model_of_price_and_weight$fitted.values
    ),
    aes(x = predicted_price, y = externally_studentized_residual)
) +
    geom_point(alpha = 0.2) +
    geom_hline(yintercept = 0, color = "red") +
    labs(
        x = "predicted price",
        y = "externally studentized residual",
        title = "Externally Studentized Residual\nvs. Predicted Price"
    ) +
theme(
    plot.title = element_text(hjust = 0.5, size = 11),
    axis.text.x = element_text(angle = 0)
)
```

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
acf(linear_model_of_price_and_weight$residuals, main = "ACF Value vs. Lag for Transformed Linear Model")
```

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
qqnorm(linear_model_of_price_and_weight$residuals)
qqline(linear_model_of_price_and_weight$residuals, col = "red")
```

# Simple Linear Regression of Price versus Weight

Simple linear regression depends of the following five assumptions being met.

1. The relationship between response / price and predictor / regressor / weight is linear, at least approximately. This assumption is not met.
   The relationship appears to be nonlinear.
2. The residuals of the linear model of price versus weight have mean $0$. This assumption is not met.
   observations are not scattered evenly around the fitted line.
   Residuals are not evenly scattered around $e = 0$.
3. The distributions of residuals of the linear model of price versus weight for different weights have constant variance.
   This assumption is not met. The vertical variation of observations is not constant. Residuals are not evenly scattered around $e = 0$.
4. The residuals of the linear model of price versus weight are uncorrelated. This assumption is not met.
   The ACF value for lag $0$ is always $1$; the correlation of the vector of residuals with itself is always 1.
   Since the ACF value for lag $3$ is significant, we have sufficient evidence to reject a null hypothesis that
   the residuals of the linear model of price versus weight are uncorrelated.
   We have sufficient evidence to conclude that the residuals of the linear model of price versus weight are correlated.
   We have sufficient evidence to conclude that the assumption that the residuals are uncorrelated is not met.
5. The residuals of the linear model of price versus weight are normally distributed. This assumption is not met.
   A linear model is robust to these assumptions.
   Given a moderate downward and a sharp upward curve at extremes of a plot of sample quantiles versus theoretical quantiles
   for the residuals of the linear model of price versus weight,
   the tails of a probability vs. externally studentized residuals plot / distribution are too light.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE}
library(TomLeversRPackage)
linear_model <- lm(price ~ weight, data = data_set)
```

Given that the above assumptions for simple linear regression are not met, we generate a data set of transformed price and/or
transformed weight such that all of the assumptions are met.
The Box-Cox Method is presented in section 5.4.1 of *Introduction to Linear Regression Analysis* (Sixth Edition)
by Douglas C. Montgomery et al..
To generate a data set of transformed price and/or transformed weight for which
the assumption that the relationship between transformed price and/or transformed weight is linear is met, and
the assumption that the distributions of residuals of a linear model of transformed price and/or transformed weight
for different weights or transformed weights have constant variance,
using R, we perform the Box-Cox Method to determine a maximum-likelihood estimate of a parameter $\lambda = 0.311$
to be used in a power transformation $y' = y^\lambda$ of a price $y$.
The maximum likelihood estimate of $\lambda$ is close to a whole parameter $\lambda = 0$.
Given a power-transformation parameter $\lambda = 0$, We transform price values according to $y' = ln(y)$.
We present a scatterplot of logarithmicized price versus weight.
We construct a linear model of logarithmicized price versus weight and present a scatterplot of externally studentized residuals
versus predicted logarithmicized price for the linear model of logarithmicized price versus weight.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
result_of_Box_Cox_Method <- perform_Box_Cox_Method(linear_model, vector_of_values_of_lambda = seq(0.29, 0.33, 0.01), whether_to_plot = FALSE)
maximum_likelihood_estimate_of_parameter_lambda <- result_of_Box_Cox_Method$maximum_likelihood_estimate_of_parameter_lambda

data_set <- data_set %>% mutate(logarithmicized_price = log(price))
ggplot() +
    geom_point(data = data_set, aes(x = weight, y = logarithmicized_price), alpha = 0.2) +
    geom_smooth(data = data_set, aes(x = weight, y = logarithmicized_price), method = "lm", se = FALSE) +
    labs(
        x = "weight (carat)",
        y = "logarithmicized price",
        title = "Logarithmicized Price vs. Weight"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )

linear_model_of_logarithmicized_price_and_weight <- lm(logarithmicized_price ~ weight, data = data_set)

ggplot(
    data.frame(
        externally_studentized_residual = linear_model_of_logarithmicized_price_and_weight$residuals,
        predicted_logarithmicized_price = linear_model_of_logarithmicized_price_and_weight$fitted.values
    ),
    aes(x = predicted_logarithmicized_price, y = externally_studentized_residual)
) +
    geom_point(alpha = 0.2) +
    geom_hline(yintercept = 0, color = "red") +
    labs(
        x = "predicted logarithmicized price",
        y = "externally studentized residual",
        title = "Externally Studentized Residual\nvs. Predicted Logarithmicized Price"
    ) +
theme(
    plot.title = element_text(hjust = 0.5, size = 11),
    axis.text.x = element_text(angle = 0)
)
```

We have generated a data set of logarithmicized price versus weight for which the assumption that
the distributions of residuals of a linear model of logarithmicized price and/or weight for different weights have constant variance.
To generate a data set of transformed price and/or transformed weight for which
the assumption that the relationship between transformed price and/or transformed weight is linear is met, and
the assumption that the mean of the residuals of the linear model of transformed price and/or transformed weight is $0$,
we take our logarithmicized-price data and additionally logarithmicize weight.
We present a scatterplot of logarithmicized price versus logarithmicized weight.
We construct a linear model of logarithmicized price versus logarithmicized weight.
We present a scatterplot of externally studentized residuals versus predicted logarithmicized price for the linear model of
logarithmicized price versus logarithmicized weight.
Additionally, we construct a plot of AutoCorrelation Function values versus lag for the linear model, and
a plot of sample quantiles versus theoretical quantiles for the residuals of the model.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width=3.3, fig.height=2.4}
data_set <- data_set %>% mutate(logarithmicized_weight = log(weight))
ggplot() +
    geom_point(data = data_set, aes(x = logarithmicized_weight, y = logarithmicized_price), alpha = 0.2) +
    geom_smooth(data = data_set, aes(x = logarithmicized_weight, y = logarithmicized_price), method = "lm", se = FALSE) +
    labs(
        x = "logarithmicized weight",
        y = "logarithmicized price",
        title = "Logarithmicized Price\nvs. Logarithmicized Weight"
    ) +
    theme(
        plot.title = element_text(hjust = 0.5, size = 11),
        axis.text.x = element_text(angle = 0)
    )

linear_model_of_logarithmicized_price_and_logarithmicized_weight <- lm(logarithmicized_price ~ logarithmicized_weight, data = data_set)
#print(summarize_linear_model(linear_model_of_logarithmicized_price_and_logarithmicized_weight))

ggplot(
    data.frame(
        externally_studentized_residual = linear_model_of_logarithmicized_price_and_logarithmicized_weight$residuals,
        predicted_logarithmicized_price = linear_model_of_logarithmicized_price_and_logarithmicized_weight$fitted.values
    ),
    aes(x = predicted_logarithmicized_price, y = externally_studentized_residual)
) +
    geom_point(alpha = 0.2) +
    geom_hline(yintercept = 0, color = "red") +
    labs(
        x = "predicted logarithmicized price",
        y = "externally studentized residual",
        title = "Externally Studentized Residual\nvs. Predicted Logarithmicized Price"
    ) +
theme(
    plot.title = element_text(hjust = 0.5, size = 11),
    axis.text.x = element_text(angle = 0)
)
```

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
acf(linear_model_of_logarithmicized_price_and_weight$residuals, main = "ACF Value vs. Lag for Transformed Linear Model")
```

1. The assumption that the relationship between response / logarithmicized price and predictor / regressor / logarithmicized weight is linear,
   at least approximately, is met.
   The relationship appears to be linear.
2. The assumption that the residuals of the linear model of logarithmicized price versus logarithmicized weight have mean $0$ is met.
   observations are scattered evenly around the fitted line.
   Residuals are evenly scattered around $e = 0$.
3. The assumptions that the distributions of residuals of the linear model of logarithmicized price versus logarithmized weight
   for different weights have constant variance is met.
   The vertical variation of observations is constant.
   Residuals are evenly scattered around $e = 0$.
4. The assumption that the residuals of the linear model of logarithmicized price versus logarithmicized weight are uncorrelated is met.
   The ACF value for lag $0$ is always $1$; the correlation of the vector of residuals with itself is always 1.
   Since all ACF value are insignificant, we have insufficient evidence to reject a null hypothesis that
   the residuals of the linear model of logarithmicized price versus logarithmicized weight are uncorrelated.
   We have insufficient evidence to conclude that the residuals of the linear model of logarithmicized price versus logarithmicized weight
   are correlated.
   We have insufficient evidence to conclude that the assumption that the residuals of the linear model of logarithmicized price
   versus logarithmicized weight are uncorrelated is not met.
5. The assumption that the residuals of the linear model of logarithmicized price versus logarithmicized weight are normally distributed
   is not met.
   However, a linear model is robust to these assumptions.
   Given a sharp downward curve at the bottom left of a plot of sample quantiles versus theoretical quantiles
   for the residuals of the linear model of logarithmicized price versus logarithmicized weight,
   a probability vs. externally studentized residuals plot / distribution is not normal.

```{r, eval=TRUE, echo=FALSE, results="show", message=FALSE, warning = FALSE, fig.width = 5.3, fig.height = 3.8, fig.align='center'}
qqnorm(linear_model_of_logarithmicized_price_and_weight$residuals)
qqline(linear_model_of_logarithmicized_price_and_weight$residuals, col = "red")
```

We determine an estimated linear-regression equation
$$\hat{\beta}_0 = 8.521$$
$$\hat{\beta}_1 = 1.944$$
$$ln(\hat{y}) = \hat{\beta}_0 + \hat{\beta}_1 ln(x)$$
where $\hat{y} = E(y|x)$ is the expected price given a weight $x$.
$$\hat{y} = exp[\hat{\beta}_0 + \hat{\beta}_1 ln(x)]$$
Consider weights $x$ and $x_+ = (1 + p) x$ and corresponding predicted prices $\hat{y}$ and $\hat{y}_+$.
$$ln(\hat{y}) = \hat{\beta}_0 + \hat{\beta}_1 ln(x)$$
$$ln\left(\hat{y}_+\right) = \hat{\beta}_0 + \hat{\beta}_1 ln\left(x_+\right)$$
$$ln\left(\hat{y}_+\right) - ln\left(\hat{y}\right) = \left[\hat{\beta}_0 + \hat{\beta}_1 ln\left(x_+\right)\right] - \left[\hat{\beta}_0 + \hat{\beta}_1 ln\left(x\right)\right]$$
$$ln\left(\frac{\hat{y}_+}{\hat{y}}\right) = \hat{\beta}_1 ln\left(x_+\right) - \hat{\beta}_1 ln\left(x\right)$$
$$ln\left(\frac{\hat{y}_+}{\hat{y}}\right) = \hat{\beta}_1 \left[ln\left(x_+\right) - ln\left(x\right)\right]$$
$$ln\left(\frac{\hat{y}_+}{\hat{y}}\right) = \hat{\beta}_1 ln\left(\frac{x_+}{x}\right)$$
$$ln\left(\frac{\hat{y}_+}{\hat{y}}\right) = \hat{\beta}_1 ln\left(\frac{(1 + p)x}{x}\right)$$
$$ln\left(\frac{\hat{y}_+}{\hat{y}}\right) = \hat{\beta}_1 ln\left(1 + p\right)$$
$$ln\left(\frac{\hat{y}_+}{\hat{y}}\right) = ln\left[\left(1 + p\right)^{\hat{\beta}_1}\right]$$
$$\frac{\hat{y}_+}{\hat{y}} = \left(1 + p\right)^{\hat{\beta}_1}$$
For an increase in weight by proportion $p$, the predicted price increases by a factor of $(1 + p)^{\hat{\beta}_1}$. For example, for an increase in weight by proportion $0.1$, or $10$ percent, the predicted price increases
by a factor of $(1 + 0.1)^{1.944} = 1.2$.
For an increase in weight from $1 \ carat$ to $1.1 \ carat$,
price increases from $exp\left[8.521 + 1.944 \ ln(1)\right] \ dollars = \$5,019.07$
by a factor of $1.2$ to $exp\left[8.521 + 1.944 \ ln(1.1)\right] \ dollars = \$6,040.75$. These prices have a ratio of $1.2$.
