---
title: "DS-6030 Homework Module 10"
author: "Tom Lever"
date: 07/31/2023
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

<!--- Below are global settings for knitr. You can override any of them by adding the changes to individual chunks --->

```{r global_options, include = FALSE}
# This chunk is called global_options. Due to `include = FALSE`, when the document is rendered, the chunk will be executed, but the results and code will not be included in the rendered document
knitr::opts_chunk$set(
    collapse = FALSE, # code and corresponding output appear in knit file in separate blocks
    echo = TRUE, # echo code by default
    comment = "#", # change comment character
    #fig.width = 5.5, # set figure width
    fig.align = "center", # set figure position
    #out.width = "49%", # set width of displayed images
    warning = TRUE, # do not show R warnings
    message = TRUE # do not show R messages
)
```

**DS 6030 \| Spring 2023 \| University of Virginia**


8.  In Section 12.2.3, a formula for calculating Proportion of Variance Explained (PVE) was given in Equation 12.10. We also saw that the PVE can be obtained using the `sdev` output of the `prcomp()` function.

    On the `USArrests` data, calculate PVE in the following two ways. These two approaches should give the same results.

    Hint: You will only obtain the same results in (a) and (b) if the same data is used in both cases. For instance, if in (a) you performed `prcomp()` using centered and scaled variables, then you must center and scale the variables before applying Equation 10.3 in (b).

    (a) Using the `sdev` output of the `prcomp()` function, as was done in Section 12.2.3.
    
        ```{r}
        the_prcomp <- prcomp(USArrests, scale = TRUE)
        variances_of_principal_components <- the_prcomp$sdev^2
        vector_of_PVEs <-
            variances_of_principal_components / sum(variances_of_principal_components)
        vector_of_PVEs
        ```

    (b) By applying Equation 12.10 directly. That is, use the `prcomp()` function to compute the principal component loadings. Then, use those loadings in Equation 12.10 to obtain the PVE.
    
        ```{r}
        matrix_of_variable_loadings <- the_prcomp$rotation
        centered_and_scaled_USArrests <- scale(USArrests)
        centered_and_scaled_USArrests_matrix <-
            as.matrix(centered_and_scaled_USArrests)
        squared_centered_and_scaled_USArrests_matrix <-
            centered_and_scaled_USArrests_matrix^2
        sum_of_squared_centered_and_scaled_USArrests_matrix <-
            sum(squared_centered_and_scaled_USArrests_matrix)
        product_of_centered_and_scaled_USArrests_matrix_and_matrix_of_variable_loadings <-
        centered_and_scaled_USArrests_matrix %*% matrix_of_variable_loadings
        product_of_centered_and_scaled_USArrests_matrix_and_matrix_of_variable_loadings_summed_on_columns <-
            apply(product_of_centered_and_scaled_USArrests_matrix_and_matrix_of_variable_loadings^2, 2, sum)
        vector_of_PVEs <-
            product_of_centered_and_scaled_USArrests_matrix_and_matrix_of_variable_loadings_summed_on_columns / 
            sum_of_squared_centered_and_scaled_USArrests_matrix
        vector_of_PVEs
        ```

9.  Consider the `USArrests` data. We will now perform hierarchical clustering on the states.

    (a) Using hierarchical clustering with complete linkage and Euclidean distance, cluster the states.

    (b) Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters?

    (c) Hierarchically cluster the states using complete linkage and Euclidean distance, after scaling the variables to have standard deviation one.

    (d) What effect does scaling the variables have on the hierarchical clustering obtained? In your opinion, should the variables be scaled before the inter-observation dissimilarities are computed? Provide a justification for your answer.