---
title: "Stat 6021: Homework Set 8"
author: "Tom Lever"
date: 10/28/22
output:
  pdf_document: default
  html_document: default
urlcolor: blue
linkcolor: red
---

1.  You will use the `birthwt` data set from the `MASS` package for this question. The data were collected at Baystate Medical Center, Springfield,
    MA in 1986. The data contain information regarding weights of newborn babies as well as potential predictors. For this question, we will focus
    on using two predictors: $age$, the mother's age in years, and $race$, the mother's race, which is coded as $1$ for white, $2$ for black, and
    $3$ for other. The response variable is `bwt`, the weight of the baby at birth in grams.
    
    (a) Produce of scatterplot of $bwt$ versus $age$.
    
        ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
        library(MASS)
        library(ggplot2)
        birthwt$race <- factor(birthwt$race)
        levels(birthwt$race) <- c("white", "black", "other")
        head(birthwt, n = 3)
        ggplot(
            birthwt,
            aes(x = age, y = bwt, color = race)
        ) +
            geom_point(alpha = 0.2) +
            geom_smooth(method = "lm", se = FALSE) +
            labs(
                x = "age",
                y = "birth weight (g)",
                title = "Birth Weight vs. Mother's Age"
            ) +
            theme(
                plot.title = element_text(hjust = 0.5),
                axis.text.x = element_text(angle = 0)
            )
        ```
        
        The slope of simple linear regression equations for birth weight versus mother's age differs for different racial categories.
        The effect on birth weight of mother's age depends on racial category.
        The influence on birth weight of mother's age differs for different racial categories.
        Because the effect on birth weight of mother's age depends on racial category, there is an interaction effect between mother's age and
        racial category.
        
    (b) Fit a regression equation with interaction between two predictors. How does this regression equation relate the age of the mother and the
        weight of the baby at birth for each of the three racial categories?
        
        ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
        library(TomLeversRPackage)
        generate_data_frame_for_indicator_variables(birthwt$race)
        linear_model <- lm(bwt ~ age * race, birthwt)
        summarize_linear_model(linear_model)
        ```
        
        The below multiple linear regression model and equation relates birthweight to mother's age and racial category.
        When a mother is white, $I_{black}$ and $I_{other}$ are $0$. When a mother is black, $I_{black}$ is $1$ and $I_{other}$ is $0$.
        When a mother has a racial category of other, $I_{black}$ is $0$ and $I_{other}$ is $1$.
        
        $$\hat{\beta}_0 = 2583.539 \ g$$
        $$\hat{\beta}_{age} = 21.373 \ \frac{g}{y}$$
        $$\hat{\beta}_{black} = 1022.788 \ g$$
        $$\hat{\beta}_{other} = 326.049 \ g$$
        $$\hat{\beta}_{age, black} = -62.538 \ \frac{g}{y}$$
        $$\hat{\beta}_{age, other} = -26.032 \ \frac{g}{y}$$
        $$y = \beta_0 + \beta_{age} \ age + \beta_{black} I_{black} + \beta_{other} I_{other} + \beta_{age, black} \ age \ I_{black} + \beta_{age, other} \ age \ I_{other} + \epsilon$$
        $$E(y | (age, white)) = \beta_0 + \beta_{age} \ age + \beta_{black} I_{black} + \beta_{other} I_{other} + \beta_{age, black} \ age \ I_{black} + \beta_{age, other} \ age \ I_{other}$$
        
        When a mother is white, birth weight varies with mother's age at a positive reference rate of $\beta_{age}$ with a positive reference bias of $\beta_0$.
        $$E(y | (age, white)) = \beta_0 + \beta_{age} \ age + \beta_{black} (0) + \beta_{other} (0) + \beta_{age, black} \ age \ (0) + \beta_{age, other} \ age \ (0)$$
        $$E(y | (age, white)) = \beta_0 + \beta_{age} \ age$$
        
        When a mother is black, birth weight varies with mother's age at a decreased, negative rate of $\beta_{age} + \beta_{age, black}$ with a increased, positive bias of $\beta_0 + \beta_{black}$.
        $$E(y | (age, black)) = \beta_0 + \beta_{age} \ age + \beta_{black} (1) + \beta_{other} (0) + \beta_{age, black} \ age \ (1) + \beta_{age, other} \ age \ (0)$$
        $$E(y | (age, black)) = \beta_0 + \beta_{age} \ age + \beta_{black} + \beta_{age, black} age$$
        $$E(y | (age, black)) = \left(\beta_0 + \beta_{black}\right) + \left(\beta_{age} + \beta_{age, black}\right) age$$
        
        When a mother has a racial category of other, birth weight varies with mother's age at a negative rate of $\beta_{age} + \beta_{age, other}$ between the reference rate and the rate for black mothers with a positive bias of $\beta_0 + \beta_{other}$ between the reference bias and the bias for black mothers.
        $$E(y | (age, other)) = \beta_0 + \beta_{age} \ age + \beta_{black} (0) + \beta_{other} (1) + \beta_{age, black} \ age \ (0) + \beta_{age, other} \ age \ (1)$$
        $$E(y | (age, other)) = \beta_0 + \beta_{age} \ age + \beta_{other} + \beta_{age, other} age$$
        $$E(y | (age, other)) = \left(\beta_0 + \beta_{other}\right) + \left(\beta_{age} + \beta_{age, other}\right) age$$
        
2.  You may only use R as a simple calculator or to find $p$-values or critical values. This question is based on data about teacher salaries from
    the $50$ states plus Washington, DC in the mid 1980's. The variables are
    
    * $AREA$: region and an element of the set {North, South, West}
    * $n$: number of political entities in an area
    * $PAY$, $y$: average annual public school teacher salary in dollars
    * $SPEND$, $x_1$: spending on public schools per student in dollars
    
    Table $1$ below provides some summary statistics of the data:
    
    $AREA$ $n$ $\bar{y}$ $\bar{x_1}$
    ------ --- --------- -----------
    North  21  24,424    3,901
    South  17  22,894    3,274
    West   13  26,159    3,919
    
    (a) Based only on Table $1$, briefly comment on the relationship between geographic area and mean teacher pay.
    
        Mean teacher pay increases going from area South to North to West.
        
    (b) Based only on Table $1$, briefly comment on the relationship between mean public school expenditure per student and mean teacher pay.
    
        As mean spending on public schools per students increases, average annual public school teacher salary increases, at an increasing rate.
        
    (c) Briefly explain why using a multiple linear regression model with teacher pay as the response variable with geographic area and public
        school expenditure per student can give further insight into the relationships between these variables.
        
        A multiple linear regression model with teacher pay as the response variable with geographic area and public school expenditure per 
        student would allow us to determine whether the effect on teacher pay of public school expenditure per student depends on geographic 
        area, and to determine whether there is an interaction effect between public school expenditure per student and geographic area.
        
3.  You may only use R as a simple calculator or to find $p$-values or critical values. This question is a continuation of question 2. A
    full multiple linear regression model with an interaction term between public school expenditure per student and geographic area $M_{full}$ 
    was fitted: The resulting multiple linear regression equation is
    $$E(y | x) = \beta_0 + \beta_1 x_1 + \beta_2 I_2 + \beta_3 I_3 + \beta_4 x_1 I_2 + \beta_5 x_1 I_3$$
    where $I_2$ and $I_3$ are the dummy codes for $AREA$ / the indicator variables for $South$ and $West$, respectively. If $AREA = South$,
    $I_2 = 1$; else $I_2 = 0$. If $AREA = West$, $I_3 = 1$; else $I_3 = 0$. The prompt for this homework includes output for an R analysis
    of variance for $M_{full}$.
    
    (a) Carry out a hypothesis test to see if the interaction terms are significant.
    
        Let $\boldsymbol{\beta}_{predictor}$ be a column vector of the coefficients of the non-reference indicator variables associated with predictor $predictor$.
        Let $\boldsymbol{predictor}$ be a column vector of the non-reference indicator variables associated with predictor $predictor$.
        The R analysis of variance considers a version of the equation for MLR model $M_{full}$
        $$E(y | x) = \beta_0 + \beta_1 x_1 + \boldsymbol{\beta}_{area} \cdot \boldsymbol{area} + \boldsymbol{\beta}_{x_1, area} \ x_1 \cdot \boldsymbol{area}$$
        
        The mean square for the interaction term $x_1 \ area$
        $$MS_{x_1, area} = \frac{SS_{x_1, area}}{DF_{x_1, area}} = \frac{9,720,281}{2} = 4,860,140.5$$
        
        We conduct a Partial $F$ Test of a null hypothesis that that regression coefficient for the interaction term is $0$, is insignificant,
        and may be dropped from the multiple linear regression model.
        
        The $F$ statistic for the Partial $F$ Test follows a $F$ distribution with $DF_R = 2$ and $DF_{Res} = 45$ degrees of freedom.
        $$F_0 = \frac{MS_{x_1, area}}{MS_{Res}} = \frac{4,860,140.5}{5,166,633} = 0.941$$
        
        The critical value of the $F$ distribution for which the probability of a random $F$ statistic being greater is equal to a 
        significance level $\alpha = 0.05$
        $$F_c = 3.204$$
        
        ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
        F_statistic <- 0.941
        significance_level <- 0.05
        regression_degrees_of_freedom <- 2
        residual_degrees_of_freedom <- 45
        qf(
            significance_level,
            regression_degrees_of_freedom,
            residual_degrees_of_freedom,
            lower.tail = FALSE
        )
        ```
        
        Since our $F$ statistic is less than the critical value, we have insufficient evidence to reject the null hypothesis that the regression
        coefficient for the interaction term is $0$, is insignificant, and may be dropped from the multiple linear model.
        
        The probability that a random $F$ statistic is greater than our $F$ statistic
        $$p = 0.398$$
        
        ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
        pf(
            F_statistic,
            regression_degrees_of_freedom,
            residual_degrees_of_freedom,
            lower.tail = FALSE
        )
        ```
        
        Since this probability is greater than a significance level $\alpha = 0.05$, we have insufficient evidence to reject the null hypothesis
        that the regression coefficient for the interaction term is $0$, is insignificant, and may be dropped from the multiple linear model.
        
        We conclude that the regression coefficient for the interaction term is $0$, is insignificant, and may be dropped from the multiple
        linear model.
        
    (b) Regardless of your answer from part 3a, suppose the interaction terms are dropped. The prompt for this homework includes a summary and
        variance-covariance  matrix for the multiple linear model without interaction. What is the reference class for this model?
        
        Consider the multiple linear model of teacher pay versus public school expenditure per student and geographic area without an interaction
        effect between public school expenditure per student and geographic area $M_{reduced}$ with equation
        $$\hat{\beta}_0 = 11,600 \ dollars$$
        $$\hat{\beta}_1 = 3.289$$
        $$\hat{\beta}_2 = 529.4 \ dollars$$
        $$\hat{\beta}_3 = 1,674 \ dollars$$
        $$E(y | x) = \beta_0 + \beta_1 x_1 + \beta_2 I_2 + \beta_3 I_3$$
        
        Because if $AREA = South$, $I_2 = 1$ and $I_2 = 0$ otherwise, and if $AREA = West$, $I_3 = 1$ and $I_3 = 0$ otherwise, when
        $AREA = North$, $I_2 = 0$ and $I_3 = 0$ and North is the reference class for $M_{reduced}$. When $Area = North$,
        $$E_N = \beta_0 + \beta_1 x_1 + \beta_2 (0) + \beta_3 (0)$$
        $$E_N = \beta_0 + \beta_1 x_1$$
        
    (c) What is the estimate of $\beta_2$? Give an interpretation of this value.
    
        When $Area = South$, $I_2 = 1$, and
        $$E_S = \beta_0 + \beta_1 x_1 + \beta_2 (1) + \beta_3 (0)$$
        $$E_S = \beta_0 + \beta_1 x_1 + \beta_2$$
        $$E_S = \left(\beta_0 + \beta_2\right) + \beta_1 x_1$$
        
        $$E_S - E_N = \beta_2$$
        $\hat{\beta}_2 = 529.4 \ dollars$ is the estimated difference between mean teacher pay when $AREA = South$ and the reference mean teacher
        pay when $AREA = North$, per $M_{reduced}$.
        
    (d) Using the Bonferroni procedure, compute the 95-percent family confidence intervals for the difference in mean response for $PAY$ between
        teachers in the
        
        i.   North region and the South region
        ii.  North region and the West region
        iii. South region and the West region
        
        When $Area = West$, $I_3 = 1$, and
        $$E_W = \beta_0 + \beta_1 x_1 + \beta_2 (0) + \beta_3 (1)$$
        $$E_W = \beta_0 + \beta_1 x_1 + \beta_3$$
        $$E_W = \left(\beta_0 + \beta_3\right) + \beta_1 x_1$$
        
        $\hat{\beta}_3 = 1,674 \ dollars$ is the estimated difference between mean teacher pay when $AREA = West$ and the reference mean teacher 
        pay when $AREA = North$, per $M_{reduced}$.
        
        $\hat{\beta}_3 - \hat{\beta}_2 = (1,674 \ dollars) - (529.4 \ dollars) = 1144.6 \ dollars$ is the estimated difference between mean 
        teacher pay when $AREA = West$ and mean teacher pay when $AREA = South$, per $M_{reduced}$.
        
        Considering the variance-covariance matrix for $M_{reduced}$,
        $$\widehat{Var\left(\beta_2\right)} = 588,126.717 \ dollars^2$$
        $$\widehat{SE\left(\beta_2\right)} = \sqrt{\widehat{Var\left(\beta_2\right)}} = \sqrt{588,126.717} = 766.894 \ dollars$$
        $$\widehat{Var\left(\beta_3\right)} = 641,873.8 \ dollars^2$$
        $$\widehat{SE\left(\beta_3\right)} = \sqrt{\widehat{Var\left(\beta_3\right)}} = \sqrt{641,873.8} = 801.170 \ dollars$$
        $$\widehat{Var\left(\beta_3 - \beta_2\right)} = \widehat{Var\left(\beta_3\right)} + \widehat{Var\left(\beta_2\right)} - 2 \widehat{Cov\left(\beta_3, \beta_2\right)}$$
        $$\widehat{Var\left(\beta_3 - \beta_2\right)} = \left(641,873.8 \ dollars^2\right) + \left(588,126.717 \ dollars^2\right) - 2 \left(244,238.02959 \ dollars^2\right)$$
        $$\widehat{Var\left(\beta_3 - \beta_2\right)} = \left(641,873.8 \ dollars^2\right) + \left(588,126.717 \ dollars^2\right) - 2 \left(244,238.02959 \ dollars^2\right)$$
        $$\widehat{Var\left(\beta_3 - \beta_2\right)} = 741,524.458 \ dollars^2$$
        $$\widehat{SE\left(\beta_3 - \beta_2\right)} = \widehat{SE\left(d_{W, S}\right)} = \sqrt{\widehat{Var\left(\beta_3 - \beta_2\right)}} = \sqrt{741,524.458} = 861.118 \ dollars$$
        
        We are making three pairwise comparisons / calculating three differences in mean response for teacher $PAY$; we consider a family of
        $g = 3$ confidence intervals.
        
        The residual degrees of freedom $DF_{Res, full} = 45$.
        When the interaction term is removed from $M_{full}$, regression degrees of freedom decreases by $2$ and residual degrees of freedom
        increases by $2$; the residual degrees of freedom $DF_{Res, reduced} = 47$.
        
        The critical value for each confidence interval
        $$t_c = t_{\alpha/(2g) = 0.05/(2*3), DF_{Res, reduced} = 47} = 2.483$$
        
        ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
        number_of_confidence_intervals <- 3
        residual_degrees_of_freedom <- 47
        qt(
            significance_level/(2*number_of_confidence_intervals),
            residual_degrees_of_freedom,
            lower.tail = FALSE
        )
        calculate_critical_value_tc(
            significance_level,
            number_of_confidence_intervals,
            residual_degrees_of_freedom
        )
        ```
        
        Let $d_{i, j}$ be the difference in mean response for $PAY$ between teachers in region $i$ and region $j$.
        For example, $d_{W, S} = \hat{\beta}_3 - \hat{\beta}_2 = 1,144.6 \ dollars$.
        
        The 95-percent family confidence interval for the difference in mean response for $PAY$ between teachers in the North region and the 
        South region
        $$\left[-\hat{d}_{S, N} - t_c \ \widehat{SE\left(-d_{S, N}\right)}, \ -\hat{d}_{S, N} + t_c \ \widehat{SE\left(-d_{S, N}\right)}\right]$$
        $$\left[-\hat{d}_{S, N} - t_c \ \widehat{SE\left(d_{S, N}\right)}, \ -\hat{d}_{S, N} + t_c \ \widehat{SE\left(d_{S, N}\right)}\right]$$
        $$\left[-\hat{\beta}_2 - t_c \ \widehat{SE\left(\beta_2\right)}, \ -\hat{\beta}_2 + t_c \ \widehat{SE\left(\beta_2\right)}\right]$$
        $$\left[-(529.4 \ dollars) - (2.483)(766.894 \ dollars), \ -(529.4 \ dollars) + (2.483)(766.894 \ dollars)\right]$$
        $$\left[-2,433.598 \ dollars, \ 1,374.798 \ dollars\right]$$
        
        The 95-percent family confidence interval for the difference in mean response for $PAY$ between teachers in the North region and the 
        West region
        $$\left[-\hat{d}_{W, N} - t_c \ \widehat{SE\left(d_{W, N}\right)}, \ -\hat{d}_{W, N} + t_c \ \widehat{SE\left(d_{W, N}\right)}\right]$$
        $$\left[-\hat{\beta}_3 - t_c \ \widehat{SE\left(\beta_3\right)}, \ -\hat{\beta}_3 + t_c \ \widehat{SE\left(\beta_3\right)}\right]$$
        $$\left[-(1,674 \ dollars) - (2.483)(801.170 \ dollars), \ -(1,674 \ dollars) + (2.483)(801.170 \ dollars)\right]$$
        $$\left[-3,663.305 \ dollars, \ 315.305 \ dollars\right]$$
        
        The 95-percent family confidence interval for the difference in mean response for $PAY$ between teachers in the South region and the 
        West region
        $$\left[-\hat{d}_{W, S} - t_c \ \widehat{SE\left(d_{W, S}\right)}, \ -\hat{d}_{W, S} + t_c \ \widehat{SE\left(d_{W, S}\right)}\right]$$
        $$\left[-(1,144.6 \ dollars) - (2.483) (861.118 \ dollars), \ -(1,144.6 \ dollars) + (2.483) (861.118 \ dollars)\right]$$
        $$\left[-3282.756 \ dollars, \ 993.556 \ dollars\right]$$
        
    (e) What do your intervals from part 3d indicate about the effect of geographic region on mean annual salary for teachers?
    
        Because all three confidence intervals contain $0$, we have insufficient evidence to reject a null hypothesis that the difference between
        any two mean annual teacher salaries for the same public school expenditure per student and different region are $0$, per $M_{reduced}$. 
        We conclude  that the difference between any two mean annual teacher salaries are approximately $0$ and that there is an insignificant 
        difference between any two mean annual teacher salaries for the same public school expenditure per student and different region.