---
title: "Building A Model That Helps Locating Displaced People"
author: "Tom Lever"
date: 06/08/2023
output:
  pdf_document: default
  html_document: default
urlcolor: blue
---

<!--- Below are global settings for knitr. You can override any of them by adding the changes to individual chunks --->
```{r global_options, include = FALSE}
# This chunk is called global_options. Due to `include = FALSE`, when the document is rendered, the chunk will be executed, but the results and code will not be included in the rendered document
knitr::opts_chunk$set(
    error = TRUE, # Keep compiling upon error
    collapse = FALSE, # code and corresponding output appear in knit file in separate blocks
    echo = TRUE, # echo code by default
    comment = "#", # change comment character
    #fig.width = 5.5, # set figure width
    fig.align = "center", # set figure position
    #out.width = "49%", # set width of displayed images
    warning = TRUE, # do not show R warnings
    message = TRUE # do not show R messages
)
```

In this project, we build a model that would help us locate people displaced by the earthquake in Haiti in $2010$. More specifically, we build in a timely manner an accurate model that classifies pixels in geo-referenced aerial images of Haiti in $2010$ as depicting blue tarps or depicting objects that are not blue tarps. People whose homes had been destroyed by the earthquake often created temporary shelters using blue tarps. Blue tarps were good indicators of where displaced people lived.

Our training data was collected likely by applying a Region Of Interest (ROI) Tool to a high-resolution, orthorectified / geo-referenced image of Haiti in 2010. One ROI tool is described at [https://www.l3harrisgeospatial.com/docs/regionofinteresttool.html](https://www.l3harrisgeospatial.com/docs/regionofinteresttool.html). Classes may be assigned to pixels by defining Regions Of Interest.

Our training data frame consists of $63,241$ observations. Each observation consists of a class in the set $\{Vegetation, \ Soil, \ Rooftop, \ Various \ Non-Tarp, \ Blue \ Tarp\}$ and a pixel. A pixel is a colored dot. A pixel is represented by a tuple of intensities of color $Red$, $Green$, and $Blue$ in the range $0$ to $255$.

According to [https://www.esri.com/about/newsroom/insider/what-is-orthorectified-imagery/](https://www.esri.com/about/newsroom/insider/what-is-orthorectified-imagery/), an orthorectified image is an accurately georeferenced image that has been processed so that all pixels are in an accurate $(x, y)$ position on the ground. Orthorectified images have been processed to apply corrections for optical distortions from the sensor system, and apparent changes in the position of ground objects caused by the perspective of the sensor view angle and ground terrain.

We use $10$-fold cross-validation to evaluate the performance of $5$ classifiers. A classifier will classify a pixel as depicting a blue tarp or depicting an object that is not a blue tarp.

We load a data frame of classes and pixels based on an orthorectified image of Haiti at [https://www.kaggle.com/datasets/billbasener/pixel-values-from-images-over-haiti?datasetId=1899167](https://www.kaggle.com/datasets/billbasener/pixel-values-from-images-over-haiti?datasetId=1899167).

```{r}
data_frame_of_classes_and_pixels <- read.csv(
    file = 'Data_Frame_Of_Classes_And_Pixels.csv'
)
head(x = data_frame_of_classes_and_pixels, n = 3)
```

TODO: Consider 5 number summaries and means and standard deviations of Red, Green, and Blue. and across classes.

