---
title: "Stat 6021: Addressing Guided Question Set for Module 4: Inference with Simple Linear Regression"
author: "Tom Lever"
date: 09/17/22
output:
  pdf_document: default
urlcolor: blue
linkcolor: red
---

We will continue to use the dataset `penguins` from the `palmerpenguins` package. In the previous guided question set, we explored the linear relationship between body mass and flipper length of adult Gentoo penguins near Palmer Station, Antarctica.
    
1. Produce a scatterplot of body mass and flipper length for adult Gentoo penguins. Write the estimated linear regression equation.
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    library(palmerpenguins)
    library(dplyr)
    library(ggplot2)
    library(TomLeversRPackage)
    
    species_flipper_length_and_body_mass <-
        palmerpenguins::penguins %>%
            select(species, flipper_length_mm, body_mass_g) %>%
            filter(!is.na(flipper_length_mm))
    head(species_flipper_length_and_body_mass, n = 3)
    
    data_set <-
        species_flipper_length_and_body_mass %>% filter(species == "Gentoo")
    ggplot(data_set, aes(x = flipper_length_mm, y = body_mass_g)) +
        geom_point(alpha = 0.2) +
        geom_smooth(method = "lm", se = FALSE) +
        labs(
            x = "flipper length (mm)",
            y = "body mass (g)",
            title = paste(
                "Body Mass vs. Flipper Length for\n",
                "Adult Gentoo Penguins near Palmer Station",
                sep = ""
            )
        ) +
        theme(
            plot.title = element_text(hjust = 0.5),
            axis.text.x = element_text(angle = 0)
        )
    
    linear_model <- lm(body_mass_g ~ flipper_length_mm, data = data_set)
    summarize_linear_model(linear_model)
    ```
    
2. For adult Gentoo penguins near Palmer Station, what is the change in the predicted body mass (in grams)
   when flipper length increases by $1 \ mm$?
   Also report the corresponding 95-percent confidence interval for the change in the predicted body mass
   when flipper length increases by $1 \ mm$.
   
   A point estimate for the change in predicted body mass when flipper length increases by $1 \ mm$ is $54.623 \ g$,
   based on the slope $\hat{\beta}_1 = 54.623 \ \frac{g}{mm}$ of linear model of body mass vs. flipper length for adult Gentoo penguins near Palmer Station.
   
   The standard error of the statistic / estimate slope $\hat{\beta}_1$
   $$SE\left(\hat{\beta}_1\right) = \sqrt{\frac{MS_{Res}}{S_{xx}}} = \sqrt{\frac{SS_{Res} / (n-2)}{\sum_{i = 1}^n \left[ \left(x_i - \bar{x}\right)^2 \right]}} = \sqrt{\frac{\sum_{i = 1}^n \left[ {e_i}^2 \right]}{(n-2) \ \sum_{i = 1}^n \left[ \left(x_i - \bar{x}\right)^2 \right]}}$$
   $$SE\left(\hat{\beta}_1\right) = \sqrt{\frac{\sum_{i = 1}^n \left[ {\left(\hat{y}_i - \bar{y} \right)}^2 \right]}{(n-2) \ \sum_{i = 1}^n \left[ \left(x_i - \bar{x}\right)^2 \right]}}$$
   
   The confidence interval for the change in the predicted body mass when flipper length increases by $1 \ mm$ is
   the confidence interval for the slope of the linear model and is
   $$\left[\hat{\beta}_1 - t_{\alpha/2, \ df_{Res}} \ SE\left(\hat{\beta}_1\right), \ \hat{\beta}_1 + t_{\alpha/2, \ df_{Res}} \ SE\left(\hat{\beta}_1\right)\right]$$
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    slope = 54.623
    confidence_level = 0.95
    significance_level = 1 - confidence_level
    number_of_observations = 123
    number_of_parameters = 2 # x and y
    degrees_of_freedom = number_of_observations - number_of_parameters
    qt((1 + confidence_level)/2, degrees_of_freedom, lower.tail = TRUE)
    # quantile t_{\alpha/2, df_{Res}} for which probability that magnitude of random test
    # statistic is greater is half significance level
    quantile <- qt(significance_level/2, degrees_of_freedom, lower.tail = FALSE)
    quantile
    standard_error_of_slope <- 5.028
    slope - quantile * standard_error_of_slope
    slope + quantile * standard_error_of_slope
    confint(linear_model, level = confidence_level)
    ```
    
    Since the confidence interval $\left[44.668 \ \frac{g}{mm}, \ 64.577 \ \frac{g}{mm}\right]$ does not contain 0, we reject the null hypothesis $H_0: \beta_1 = 0 \ \frac{g}{mm}$.
    We have sufficient evidence to support the alternate hypothesis $H_1: \beta_1 \neq 0$.
    
    Since the confidence interval $\left[44.668 \ \frac{g}{mm}, \ 64.577 \ \frac{g}{mm}\right]$ contains $50 \ \frac{g}{mm}$, we fail to reject the null hypothesis $H_0: \beta_1 = 50 \ \frac{g}{mm}$.
    
3. Conduct a hypothesis test to determine whether or not there is a linear association between body mass and flipper length
   for adult Gentoo penguins near Palmer Station. State the hypothesis, p-value, and conclusion in context.
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    qt(confidence_level, degrees_of_freedom, lower.tail = TRUE)
    # quantile t_{\alpha, df_{Res}} for which probability that magnitude of random test
    # statistic is greater is significance level
    qt(significance_level, degrees_of_freedom, lower.tail = FALSE)
    ```
   
   Given a significance level 0.05,
   we test a null hypothesis $H_0: \beta_1 = 0$ that the slope of a linear model of body mass vs. flipper length for adult Gentoo penguins
   near Palmer station is equal to 0.
   If we have sufficient evidence to reject the null hypothesis, we have sufficient evidence to support an alternate hypothesis
   $H_1: \beta_1 \neq 0$ that the slope of the linear model is not equal to 0.
   Since the alternate hypothesis involves "$\neq$", we have sufficient evidence to reject the null hypothesis
   if the magnitude $\left|t_0\right|$ of the test statistic $t_0 = 10.863$ is greater than $t_{\alpha/2, df_{Res}} = 1.980$.
   If the alternate hypothesis were to involve "$<$" or "$>$", we would have sufficient evidence to reject the null hypothesis
   if the magnitude $\left|t_0\right|$ of the test statistic $t_0 = 10.863$ is greater than $t_{\alpha, df_{Res}} = 1.658$.
   Since $\left|t_0\right| > t_{\alpha/2, df_{Res}}$, we reject the null hypothesis. We have sufficient evidence to support the alternate hypothesis.
   $${t_0}^2 = \frac{\left(\hat{\beta}_1 - \beta_{10}\right)^2}{MS_{Res} / S_{xx}} = \frac{\left(\hat{\beta}_1 - 0\right)^2}{MS_{Res} / S_{xx}} = \frac{{\hat{\beta}_1}^2}{MS_{Res} / S_{xx}} = \frac{{\hat{\beta}_1}^2 S_{xx}}{MS_{Res}} = \frac{\hat{\beta}_1 S_{xy}}{MS_{Res}} = \frac{MS_R}{MS_{Res}} = F_0$$
   We have sufficient evidence to reject the null hypothesis if the probability $p$ that the test statistic for
   a random sample of adult Gentoo Penguins near Palmer Station is less than $-\left|t_0\right|$ or greater than $\left|t_0\right|$,
   assuming the null hypothesis is true, is less than significance level $\alpha$.
   Since $p < 2.2 \times 10^{-16}$ is less than $\alpha = 0.05$, we reject the null hypothesis. We have sufficient evidence to support the alternate hypothesis.
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    magnitude_of_test_statistic <- abs(10.863)
    probability <- pt(
        magnitude_of_test_statistic, degrees_of_freedom, lower.tail = FALSE
    ) * 2
    # 2 if the alternate hypothesis involves "\neq"
    # 1 if the alternate hypothesis involves "<" or ">"
    probability
    test_null_hypothesis_involving_slope(linear_model, 0.05)
    ```
    
4. Are your results from parts 2 and 3 consistent?

   Yes; the results of constructing the confidence interval for the slope and testing the null hypothesis support rejecting the null hypothesis
   and supporting the alternate hypothesis.
   
5. Estimate the mean body mass for adult Gentoo penguins near Palmer Station with flipper lengths of $x_0 = 200 \ mm$.
   Also report the 95-percent confidence interval for the mean body mass for Gentoo penguins with flipper lengths of $x_0$.
   
   A point estimate for the mean body mass for adult Gentoo penguins near Palmer Station with flipper lengths of $x_0 = 200 \ mm$ is
   $\hat{y} = \hat{\beta}_0 + \hat{\beta}_1 x \approx (-6787.281 \ g) + \left(54.623 \ \frac{g}{mm}\right) (200 \ mm) = 4137.22 \ g$.
   
   The standard error of the statistic / estimate mean body mass for adult Gentoo penguins near Palmer station with flipper lengths of $x_0 = 200 \ mm$
   $E\left(\hat{y} \ | \ x_0\right)$
   $$SE\left\{E\left(\hat{y} \ | \ x_0\right)\right\} = \sqrt{MS_{Res} \left[\frac{1}{n} + \frac{\left(x_0 - \bar{x}\right)^2}{S_{xx}}\right]}$$
   
   The confidence interval for the mean body mass for adult Gentoo penguins with flipper lengths of $x_0 = 200 \ mm$ is
   $$\left[ E\left(\hat{y} \ | \ x_0\right) - t_{\alpha/2, \ df_{Res}} \ SE\left\{E\left(\hat{y} \ | \ x_0\right)\right\}, \ E\left(\hat{y} \ | \ x_0\right) + t_{\alpha/2, \ df_{Res}} \ SE\left\{E\left(\hat{y} \ | \ x_0\right)\right\} \right]$$
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    predict(
        linear_model,
        data.frame(flipper_length_mm = 200),
        level = 0.95,
        interval = "confidence"
    )
    ```
    
6. Report the 95 percent prediction interval for the body mass of an adult Gentoo penguin with flipper length of $x_0 = 200 \ mm$.

   The standard error of the prediction body mass for adult Gentoo penguins near Palmer station with flipper lengths of $x_0 = 200 \ mm$
   $E\left(\hat{y} \ | \ x_0\right)$
   $$SE\left\{\hat{y} \ | \ x_0\right\} = \sqrt{MS_{Res} \left[1 + \frac{1}{n} + \frac{\left(x_0 - \bar{x}\right)^2}{S_{xx}}\right]}$$
   
   The prediction interval for the body mass for adult Gentoo penguins with flipper lengths of $x_0 = 200 \ mm$ is
   $$\left[ E\left(\hat{y} \ | \ x_0\right) - t_{\alpha/2, \ df_{Res}} \ SE\left\{\hat{y} \ | \ x_0\right\}, \ E\left(\hat{y} \ | \ x_0\right) + t_{\alpha/2, \ df_{Res}} \ SE\left\{\hat{y} \ | \ x_0\right\} \right]$$
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    predict(
        linear_model,
        data.frame(flipper_length_mm = 200),
        level = 0.95,
        interval = "prediction"
    )
    ```
    
7. A researcher hypothesizes that for adult Gentoo penguins near Palmer Station, the predicted body mass increases by more than $50 \ g$
   for each additional millimeter in flipper length. Conduct an appropriate hypothesis test. What are the null and alternate hypotheses,
   test statistic, and conclusion?
   
   Given a significance level $\alpha = 0.05$ and assumed linear-model slope $\beta_{10} = 50 \ \frac{g}{mm}$, we test a null hypothesis
   $H_0: \beta_1 \leq \beta_{10}$ that the predicted body mass increases by a mass less than or equal to $50 \ g$
   for each additional millimeter in flipper length, and that the slope of the linear model of body mass vs. flipper length
   is less than or equal to $\beta_{10}$. If we have sufficient evidence to reject the null hypothesis, we have sufficient evidence
   to support an alterate hypothesis $H_1: \beta_1 > \beta_{10}$ that the slope of the linear model
   of body mass vs. flipper length is greater than $\beta_{10}$. We have sufficient evidence to support an alternate hypothesis
   $H_1: \beta_1 > \beta_{10}$ if the magnitude $\left|t_0\right|$ of the test statistic $t_0$ is greater than $t_{\alpha, \ df_{Res}} = 1.658$, 
   where
   $$t_0 = \frac{\hat{\beta}_1 - \beta_{10}}{SE\left(\hat{\beta}_1\right)} = \frac{\left(54.623 \ \frac{g}{mm}\right) - \left(50 \ \frac{g}{mm}\right)}{\left(5.028 \ \frac{g}{mm}\right)} = 0.919$$
   Since $\left|t_0\right| = 0.919$ is less than $t_{\alpha, \ df_{Res}} = 1.658$,
   we have insufficient evidence to reject that the null hypothesis
   that the predicted body mass increases by a mass less than or equal to $50 \ g$ for each additional millimeter in flipper length,
   and that the slope of the linear model of body mass vs. flipper length is less than or equal to $\beta_{10}$.
   
   We have sufficient evidence to reject the null hypothesis if the probability $p$ that the magnitude of a test statistic for a random sample
   of adult Gentoo penguins near Palmer Station is greater than the magnitude $\left|t_0\right|$,
   assuming the null hypothesis is true, is less than significance level $\alpha$.
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    pt(0.919, degrees_of_freedom, lower.tail = FALSE) * 1
    # 2 if the alternate hypothesis involves "\neq"
    # 1 if the alternate hypothesis involves "<" or ">"
    ```
    
    Since $p = 0.180$ is greater than $\alpha$, we fail to reject the null hypothesis.
    
    If we were interested in the question, "For a single adult Gentoo penguin with a flipper length of $200 \ mm$,
    is the predicted body mass of that penguin greater than $5000 \ g$", we would test a null hypothesis $H_0: y = 5000 \ g$,
    and calculate a test statistic using $SE\left\{\hat{y} \ | \ x_0\right\}$.

    If we were interested in the question, "For adult Gentoo penguins with flipper length $200 \ mm$,
    is the average body mass greater than 5000 grams?", we would test a null hypothesis $H_0: \bar{y} = 5000 \ g$,
    and calculate a test statistic using $SE\left\{E\left(\hat{y} \ | \ x_0\right)\right\}$.
    
    In the former case, we consider the probability of observing a particular body mass;
    in the latter case, we consider the probability of observing a given mean body mass within a new sample of adult Gentoo penguins
    near Palmer Station.