---
title: "Chapter 2 Lab: Introduction to R"
# author: "**SOLUTIONS**"
#author: "**Your Name Here**"
#date: "May 16, 2020"
#output: html_document # uncomment and comment the following line to switch
output: 
  html_document:  # notebooks may be a better option for homework
    toc: true
    toc_depth: 1
    toc_float: true  
editor_options: 
  chunk_output_type: inline
---
<!--- Below are global settings for knitr. You can override any of them by adding the changes to individual chunks --->
```{r global_options, include=FALSE}
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=TRUE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.width = 5.5,     # set figure width
                      fig.align = "center",# set figure position
                      out.width = "49%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE)     # do not show R messages
```

This includes one mapping from the one-to-many mappings that exist from the base
R code presented in ILSR into a tidyverse re-implementation.
For fuller and richer exposition beyond the code artifacts present here, please 
feel free to explore other such efforts that are available online, e.g., [here](https://github.com/SmithCollege-SDS/tidy-islr/blob/master/lab01/Lab%201%20-%20Introduction%20to%20R.ipynb)
and [here](https://github.com/uc-cfss/tidy-islr/blob/master/02-stat-learn.Rmd).
In particular, these two resources have been referred to and leveraged to 
produce the tidyverse implementations we present and use below.

# Basic Commands (Base R)
## Vectors
Create a vector
```{r}
x <- c(1,3,2,5)
x
```
```{r}
x = c(1,6,2)
x
```
```{r}
y <- c(1,4,3)
```
Length of vector
```{r}
length(x)
length(y)
```
Element-wise addition of two vectors
```{r}
x + y
```
Removing objects 
```{r}
ls()
rm(x, y)
ls()
rm(list=ls()) # remove all objects in your environment
```
## Matrices
```{r}
?matrix
```
```{r}
x <- matrix(data=c(1, 2, 3, 4), nrow=2, ncol=2) # populate matrix by row
x
```
```{r}
x <- matrix(c(1, 2, 3, 4), 2, 2)  # you can ignore keywords (not recommended)
matrix(c(1, 2, 3, 4), 2, 2, byrow=TRUE)  # populate matrix by row
```
Element-wise application of functions
```{r}
sqrt(x)
x^2
```
## Random numbers
```{r}
x <- rnorm(50)
y <- x + rnorm(50, mean=50, sd=0.1)

cor(x,y)
```
Use a random seed to ensure reproducibility 
```{r}
set.seed(1303)
rnorm(50)
```
A few statistics
```{r}
set.seed(3)
y <- rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)
```


# Graphics

## Data {.tabset}
### Tidyverse

```{r}
library(tidyr)
library(tibble)

xy <- tibble(
  x=rnorm(100),
  y=rnorm(100)
)
```

### Base R
```{r}
x <- rnorm(100)
y <- rnorm(100)
```

## {-}

## Scatterplot {.tabset}
### Tidyverse
```{r scatter}
library(ggplot2)

scatter_plot <- ggplot(mapping=aes(x=x, y=y), data=xy) +
  geom_point(color="darkgreen") +
  labs(title="Plot of X vs Y",
       x="this is the x-axis",
       y="this is the y-axis")

scatter_plot
```
```
ggsave(filename="scatter_plot.pdf", plot=scatter_plot)
```

### Base R
```{r}
plot(x, y)
plot(x, y, xlab="this is the x-axis", ylab="this is the y-axis", main="Plot of X vs Y")
```
```
pdf("Figure.pdf")
  plot(x,y,col="green")
dev.off()
```

## {-}

## Contour plots {.tabset}
### Tidyverse ggplot
```{r}
d <- seq(-pi,pi,length=50)
x <- outer(d, d, function(x, y){ x })
y <- t(x)
z <- outer(d, d, function(x, y){ cos(y)/(1+x^2) })

xyz <- tibble(
  x=as.vector(x),  # x, y, and z are 2D matrices; flatten to 1D vector
  y=as.vector(y),
  z1=as.vector(z)
)

contour_plot1 <- ggplot(mapping=aes(x=x, y=y, z=z1), data=xyz) +
  geom_contour()

contour_plot1 + geom_contour(bins=45)

xyz <- xyz %>% 
  add_column(z2=as.numeric((z - t(z))/2))

contour_plot2 <- ggplot(mapping=aes(x=x, y=y, z=z2), data=xyz) +
  geom_contour(bins=15)

ggplot(mapping=aes(x=x, y=y, z=z2), data=xyz) +
  geom_contour_filled()
```

### Base R
```{r}
x=seq(1,10)
x
x=1:10
x
x=seq(-pi,pi,length=50)
y=x
f=outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f,nlevels=45,add=T)  # add to current graph!
```
```{r}
fa=(f-t(f))/2
contour(x,y,fa,nlevels=15)
```
```{r}
image(x,y,fa)
```

## 3D graphs {.tabset}
### plotly

`install.packages("plotly", repos="http://cran.rstudio.com")`

```{r plotly}
library(plotly)

plot_ly(x=x, y=y, z=z, type='surface') %>%
  layout(scene = list(aspectmode='cube'))
# zoom and drag!
```


### Base R
```{r}
persp(x,y,fa)
```
```{r}
persp(x,y,fa,theta=30)
persp(x,y,fa,theta=30,phi=20)
persp(x,y,fa,theta=30,phi=70)
persp(x,y,fa,theta=30,phi=40)
```

## {-}

# Indexing Data
```{r}
A <- matrix(1:16, 4, 4)
A
```
```{r}
A[2, 3]
```
```{r}
A[c(1,3), c(2,4)]
A[1:3, 2:4]
```
Slices
```{r}
A[1:2, ]
A[, 1:2]
A[1, ]
```
Remove 
```{r}
A[-c(1, 3),]
A[-c(1, 3), -c(1, 3, 4)]
```
```{r}
dim(A)  # the original matrix is unchanged
```

## tidyverse indexing
### dplyr: filter/select

filter: remove rows based on a condition
```{r dplyr}
library(dplyr)

filter(.data = xyz, x > 0)
xyz %>% 
  filter(x > 0)
```
select: select columns
```{r}
xyz %>%
  select(x, y)
```


# Loading Data

## Read data from file {.tabset}
### Tidyverse readr
```{r auto1}
# Load the `Auto` data using the `read_csv()` function (from `readr` package).
library(readr)
Auto <- read_csv('https://www.statlearning.com/s/Auto.csv')
```
```
View(Auto) # <-- fix(Auto), or, if you're in Rstudio and have a data object
           # you can just click on the data object
```
```{r}
Auto
Auto %>% count(horsepower)
Auto = read_csv('https://www.statlearning.com/s/Auto.csv', 
                na="?")
summary(Auto)
```
### Base R
```{r}
Auto=read.table("Auto.data")
head(Auto)
```
```{r}
Auto <- read.table("Auto.data", header=T, na.strings="?")
head(Auto)
```
```{r}
Auto <- read.csv("Auto.csv", header=T, na.strings="?")
head(Auto)
```
## Handling missing data
One way of handling missing data
```{r}
Auto[rowSums(is.na(Auto)) > 0,]
```

```{r}
dim(Auto)
Auto <- na.omit(Auto)
dim(Auto)
names(Auto)
```

The tidyverse can handle NA's appropriately without us explicitly dropping those rows

```{r, echo=FALSE}
baseAuto <- Auto
```

# Additional Graphical and Numerical Summaries
## Scatterplots and boxplots {.tabset}
### Tidyverse ggplot
```{r}
# 
ggplot(Auto, aes(x=factor(cylinders), y=mpg)) + 
  geom_boxplot(fill="orange", color="blue") + 
  labs(x="cylinders", y="MPG")

Auto <- Auto %>% 
  mutate(cylinders=as.factor(cylinders))

ggplot(data=Auto, mapping=aes(x=cylinders, y=mpg)) +
  geom_boxplot(fill="orange", color="blue") + 
  labs(x="cylinders", y="MPG")

ggplot(Auto, aes(x=mpg)) + 
  geom_histogram(binwidth=1, center=10,  
                 fill="orange", color="blue")
```

### Base R
```{r, echo=FALSE}
Auto <- baseAuto 
```

```
plot(cylinders, mpg)
```
You need to specify the data frame explicitly
```{r}
plot(Auto$cylinders, Auto$mpg)
```
Attaching a data frame makes the columns directly available. **I do not recommend this approach as it decouples the columns from the data frame.**
```{r}
attach(Auto)
plot(cylinders, mpg)
```
Treat cylinders as a qualitative variable by converting to a factor. This will switch the plot to a boxplot!
```{r}
cylinders <- as.factor(cylinders)
plot(cylinders, mpg)
```
**The column `cylinders` in the `Auto` data frame is still a numerical column and not a factor!**

Modifying the output
```{r}
plot(cylinders, mpg, col="red")
plot(cylinders, mpg, col="red", varwidth=T)  # width proportional to sqrt of sample size
plot(cylinders, mpg, col="red", varwidth=T, horizontal=T)
plot(cylinders, mpg, col="red", varwidth=T, xlab="cylinders", ylab="MPG")
```
```{r}
Auto
```
```{r, echo=FALSE}
baseAuto <- Auto
```

## {-}

## Histograms {.tabset}
### Tidyverse ggplot
```{r}
ggplot(Auto, aes(x=mpg)) + 
  geom_histogram(binwidth=1, 
                 center=10, 
                 fill="orange", 
                 color="blue")
```

### Base R
```{r}
hist(mpg)
hist(mpg,col=2)
hist(mpg,col=2,breaks=15)
```

## {-}

## Scatterplot matrices {.tabset}
### Tidyverse GGally
```{r}
library(GGally)
ggpairs(data %>% select(-name))  # remove name column
```

### Base R
```
pairs(Auto)
```
Column `name` causes a problem 
```{r}
pairs(~ mpg + displacement + horsepower + weight + acceleration, Auto)
```

## Interactive graphs {.tabset}

### plotly
```{r auto4}
library(plotly)

p <- ggplot(Auto, aes(x=horsepower, y=mpg, label=name)) +
  geom_point()
ggplotly(p)
```

### Base R
```{r}
plot(horsepower,mpg)
```
```
identify(horsepower,mpg,name)
```

## Summary

```{r}
summary(Auto)
```
```{r}
summary(mpg)
```
