---
title: "Stat 6021: Guided Question Set 6"
author: "Tom Lever"
date: 10/09/22
output:
  pdf_document: default
  html_document: default
urlcolor: blue
linkcolor: red
---

For this guided question set, we will use the data set `nfl.txt`, which contains data on NFL team performance from the 1976 season.
The variables are:

* $y$: Games won in the 14-game 1976 season
* $x_1$: Rushing yards
* $x_2$: Passing yards
* $x_3$: Punting average (yards / punt)
* $x_4$: Field-goal percentage (field goals made / field goals attempted)
* $x_5$: Turnover differential (turnovers acquired - turnovers lost)
* $x_6$: Penalty yards
* $x_7$: Percent rushing (rushing plays / total plays)
* $x_8$: Opponents' rushing yards
* $x_9$: Opponents' passing yards

1.  Create a scatterplot matrix and find the correlation between all pairs of variables for this data set.

    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    library(TomLeversRPackage)
    data_set <- read.table("nfl.txt", header = TRUE)
    head(data_set, n = 3)
    nrow(data_set)
    pairs(data_set, lower.panel = NULL)
    correlation_matrix <- round(cor(data_set), 3)
    correlation_matrix
    #linear_model <- lm(y ~ x1, data = data_set)
    #summarize_linear_model(linear_model)
    ```

    Answer the following questions based on the output.
    (a) Which predictors appear to be linearly related to the number of wins?
        Which predictors do not appear to have a linear relationship with the number of wins?
        
        We note that predictors in the set {$x_1$, $x_2$, $x_5$, $x_7$, $x_8$} have moderate to high correlations with the number of wins.
        The last predictor is the only one moderate to highly negatively associated with the number of wins.
        
        The predictors in the set {$x_3$, $x_4$, $x_6$, $x_9$} do not have a strong linear relationship with the number of wins.
        
        Considering statistics for each of nine linear models of number of games won $y$ versus one predictor $x$,

        predictor $x$ test statistic $t_0$ probability $p$
        ------------- -------------------- ---------------
        $x_1$                        3.758         0.00877
        $x_2$                        2.811         0.00927
        $x_3$                       -0.413         0.68300
        $x_4$                        1.364         0.18400
        $x_5$                        3.049         0.00522
        $x_6$                        1.172         0.25200
        $x_7$                        3.317         0.00269
        $x_8$                       -5.577        7.38e-06
        $x_9$                       -1.626         0.11609
        
        Let significance level $\alpha = 0.05$.
        
        Let $t_{\alpha/2, \ n - p} = t_{0.05/2, \ 28 - 2}$ be the quantile / critical value of a Student's $t$ distribution
        with $n - p = 28 - 2 = 26$ degrees of freedom for which the probability that a test statistic is greater is $\alpha/2 = 0.05/2 = 0.025$.
        
        Since the probability $p$, that the magnitude $\left|t\right|$ of a random test statistic following the above distribution is greater
        than $t_0$, is less than $\alpha$ for each predictor in the set {$x_1$, $x_2$, $x_5$, $x_7$, and $x_8$},
        we have sufficient evidence to reject a null hypothesis that the predictor is not linearly related to the number of wins.
        We have sufficient evidence to support an alternate hypothesis that the predictor is linearly related to the number of wins.
        Since the probability is greater than $\alpha$ for each predictor in the set {$x_3$, $x_4$, $x_6$, and $x_9$},
        we have insufficient evidence to reject a null hypothesis that the predictor is not linearly related to the number of wins.
        The predictor may not be linearly related to the number of wins.
        
    (b) Do you notice if any of the predictors are highly correlated with one another?
        If so, which ones?
        
        According to Keith G. Calkins 
        ([https://www.andrews.edu/~calkins/math/edrm611/edrm05.htm](https://www.andrews.edu/~calkins/math/edrm611/edrm05.htm)),
        "Correlation coefficients whose magnitude[s] are between 0.9 and 1.0 indicate variables which can be considered very highly correlated.
        Correlation coefficients whose magnitude[s] are between 0.7 and 0.9 indicate variables which can be considered highly correlated.
        Correlation coefficients whose magnitude[s] are between $0.5$ and $0.7$ indicate variables which can be considered moderately correlated.
        Correlation coefficients whose magnitude[s] are between $0.3$ and $0.5$ indicate variables which have... low correlation[s].
        Correlation coefficients whose magnitude[s] are less than $0.3$ have little if any (linear) correlation."
        
        ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
        ifelse(abs(correlation_matrix) > 0.7, correlation_matrix, "")
        ifelse(abs(correlation_matrix) > 0.5, correlation_matrix, "")
        ```
        
        Each predictor is highly correlated with itself.
        Rushing yards $x_1$ and percent rushing $x_7$ are highly correlated.
        $x_1$ is moderately correlated with $x_5 and $x_8$.
        $x_5$ is moderately correlated with $x_7$.
        $x_7$ is moderately correlated with $x_8$.
        
    (c) What predictors would you first consider to use in a multiple linear regression? Briefly explain your choices.
    
        I would consider using $x_1$, $x_2$, $x_5$, $x_7$, and $x_8$ in a multiple linear regression as these predictors
        have high correlation with the number of wins.
        
        I would consider using $x_2$ as it has a high correlation with the number of wins, and a subset of {$x_1$, $x_5$, $x_7$, $x_8$}
        as these are correlated with the number of wins but are also correlated among themselves.
        
2.  Regardless of your answer to the previous question, fit a multiple regression model for the number of games won against the followning three      predictors: the team's passing yardage ($x_2$), the percentage of rushing plays ($x_7$), and the opponents' yards rushing ($x_8$).
    Write the estimated regression equation.
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    linear_model <- lm(y ~ x2 + x7 + x8, data = data_set)
    summarize_linear_model(linear_model)
    ```
    
3.  Interpret the estimated coefficient for the predictor percentage of rushing plays $x_7$ in context.

    Holding predictor team's passing yardage $x_2$ and opponents' yards rushing $x_8$ constant, for an increase in percentage of rushing plays
    $x_7$ of $1 / 0.194 = 5.156$, expected games won in the 14-game 1976 season increases by $1$.
   
4.  What is the estimated number of games a team would win for a team's passing yardage $x_2 = 2000 \ yards$, percentage of rushing plays $x_7 = 48$,
    and opponents' yards rushing $x_8 = 2350 \ yards$? Also provide a relevant 95-percent confidence interval for the number of games.
    
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    linear_model <- lm(y ~ x2 + x7 + x8, data = data_set)
    predict(
        linear_model,
        newdata = data.frame(x2 = 2000, x7 = 48, x8 = 2350),
        interval = "confidence"
    )
    predict(
        linear_model,
        newdata = data.frame(x2 = 2000, x7 = 48, x8 = 2350),
        interval = "predict"
    )
    ```
    
    For a particular predictor $\vec{x}_0 = (2000 \ yards, 48, 2350 \ yards)$ in the three-dimensional space with
    $x_2^{min} \leq x_2 \leq x_2^{max}$, $x_7^{min} \leq x_7 \leq x_7^{max}$, and $x_8^{min} \leq x_8 \leq x_8^{max}$,
    
    The estimated number of a games a team would win is $3$.
    
    The $95$-percent confidence interval for the expected / mean number of games a team would win
    $E\left(y | \vec{x}_0\right)$ is $[1.711, \ 5.052]$.
    The $95$-percent prediction interval for a future number of games a team will win $y_{\vec{x}_0}$ is $[-0.516, \ 7.279]$.
    We are asked for the prediction interval.
    
5.  Using the output for the multiple linear regression model from part 2, answer the following question from a client:
    "Is this regression model useful in predicting the number of wins during the 1976 season?".
    Be sure to write the null and alternative hypotheses, state the value of the test statistic, state the $p$-value, and state a relevant
    conclusion.
    What is the critical value associated with this hypothesis test? Perform the test with significance level $\alpha = 0.05$.
    
    We assume that errors are random, are independent, and follow a normal distribution with mean $E\left(\epsilon_i\right) = 0$ and
    variance $Var\left(\epsilon_i\right) = \sigma^2$.
    The multiple linear regression model from part 2 is useful in predicting the number of wins during the 1976 season
    if at least one of the predictor variables in the set $\left\{x_2, x_7, x_8\right\}$ contributes significantly
    to the model.
    We conduct a test of the null hypothesis $H_0: \beta_{x_2} = \beta_{x_7} = \beta_{x_8} = 0$ that
    all coefficients in the set $\left\{\beta_{x_2}, \beta_{x_7}, \beta{x_8}\right\}$ are 0.
    The alternate hypothesis is $H_1: \beta_{x_2} \neq 0 \ or \ \beta_{x_7} \neq 0 \ or \ \beta_{x_8} \neq 0$ that
    at least one of the coefficients in the set $\left\{\beta_{x_2}, \beta_{x_7}, \beta{x_8}\right\}$ is not 0.
    The alternate hypothesis is also $H_1: \beta_{x_i} \neq 0 \ for \ i \in [2, 7, 8]$.
    If we reject the null hypothesis, at least one of the predictor variables contributes significantly to the model.
    
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    analyze_variance(linear_model)
    test_null_hypothesis_involving_MLR_coefficients(linear_model, 0.05)
    ```
    
    Alternatively, since the test statistic $F_0 = 29.437$ is greater than the critical value
    $F_{\alpha = 0.05, \ df_R = k = 3, \ df_{Res} = n - p = 28 - 4} = 3.009$, we reject the null hypothesis and support the alternate hypothesis.
    
    Since we reject the null hypothesis and support the alternate hypothesis, at least one of the predictor variables contributes
    significantly to the model. Since at least one of the predictor variables contributes significantly to the model,
    the model is useful in predicting the number of wins during the 1976 season.
    
6.  Report the value of the $t$ statistic for the predictor $x_7$. What is the relevant conclusion from this $t$ statistic?
    Also report the critical value for this hypothesis test. Perform the test at significance level $0.05$.
    
    The test statistic for the predictor percent rushing $x_7$ $t_0 = 2.198$. The critical value $t_{\alpha/2 = 0.05/2, \ n-p = 28-4} = 2.064$.
    
    In parallel, the probability, that the magnitude $\left|t\right|$ of a random test statistic is greater than
    the magnitude $\left|t_0\right|$ of our test statistic, $p = 0.0378$.
    
    Since the test statistic for the predictor $x_7$ is greater than the critical value, and
    the probability is less than the significance level $\alpha = 0.05$, we reject a null hypothesis that
    the individual predictor $x_7$ is insignificant in predicting the response $y$ and can be removed from the model.
    We have sufficient evidence to support the alternate hypothesis that the individual predictor percent rushing $x_7$
    is significant in predicting the response number of games won $y$ and cannot be removed from the model.
    
7. Check the regression assumptions by creating a residual plot, an ACF plot of the residuals, and a QQ plot of the residuals.
   Comment on these plots.

    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    library(ggplot2)
    ggplot(
        data.frame(
            externally_studentized_residual = linear_model$residuals,
            predicted_games_won = linear_model$fitted.values
        ),
        aes(x = predicted_games_won, y = externally_studentized_residual)
    ) +
        geom_point(alpha = 0.2) +
        geom_hline(yintercept = 0, color = "red") +
        labs(
            x = "predicted games won",
            y = "externally studentized residual",
            title = "Residuals vs. Predicted Games Won"
        ) +
        theme(
            plot.title = element_text(hjust = 0.5),
            axis.text.x = element_text(angle = 0)
        )
    
    acf(linear_model$residuals, main = "ACF Value vs. Lag for Linear Model")
    
    qqnorm(linear_model$residuals)
    qqline(linear_model$residuals, col = "red")
    ```
    
    1.  The assumption that the relationship between response / games won and predictors is linear, at least approximately, is met
        cannot be addressed.
    2.  The assumption that the residuals of the linear model of games won versus predictors have mean $0$ is met.
        Residuals are evenly scattered around $e = 0$ at random.
    3.  The assumption that the distributions of residuals of the linear model for different weights have constant variance is met.
        Residuals are evenly scattered around $e = 0$ with constant vertical variance.
    4.  The assumption that the residuals of the linear model are uncorrelated is met.
        The ACF value for lag $0$ is always $1$; the correlation of the vector of residuals with itself is always 1.
        Since all ACF value are insignificant, we have insufficient evidence to reject a null hypothesis that
        the residuals of the linear model are uncorrelated.
        We have insufficient evidence to conclude that the residuals of the linear model are correlated.
        We have insufficient evidence to conclude that the assumption that the residuals of the linear model are uncorrelated is not met.
    5.  The assumption that the residuals of the linear model are normally distributed is met.
        A linear model is robust to these assumptions.
        Considering a plot of sample quantiles versus theoretical quantiles
        for the residuals of the linear model, since observations lie near the line of best fit / their theoretical values,
        a probability vs. residuals plot / distribution is normal.
        
8. Consider adding another predictor $x_1$, the team's rushing yards for the season, to the linear model defined in part 2.
   Interpret the results of the $t$ test for the coefficient of this predictor. A classmate says: "Since the result of the $t$ test
   is insignificant, the team's rushing yards for the season is not linearly related to the number of wins".
   
    The test statistic for the predictor percent rushing $x_1$ $t_0 = 0.549$.
    The critical value $t_{\alpha/2 = 0.05/2, \ n - p = 28 - 5} = 2.069$.
   
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    linear_model <- lm(y ~ x1 + x2 + x7 + x8, data = data_set)
    summarize_linear_model(linear_model)
    ```
    
    In parallel, the probability, that the magnitude $\left|t\right|$ of a random test statistic is greater than
    the magnitude $\left|t_0\right|$ of our test statistic, $p = 0.589$.
    
    Since the test statistic for the predictor $x_1$ is less than the critical value, and
    the probability is greater than the significance level $\alpha = 0.05$, we have insufficient evidence to reject a null hypothesis that
    the individual predictor $x_1$ is insignificant in predicting the response number of games won $y$ in the presence of the other predictors.
    We have insufficient evidence to support the alternate hypothesis that the individual predictor rushing yards $x_1$
    is significant in predicting the response number of games won $y$ in the presence of the other predictors.
    I agree with my classmate in that the result of the $t$ test is insignificant.
    I would change my classmate's wording to "we have insufficient evidence to support the alternate hypothesis that the individual predictor
    rushing yards $x_1$ is significant in predicting the response number of games won $y$ in the presence of the other predictors,
    and may be removed from the model".
    
    To address the classmate's statement, we need to fit a simple linear regression model with $x_1$, the team's rushing yards,
    as the only predictor. The multiple linear regression model is not meant to address the classmate's statement.
    
    Fitting a simple linear regression model with $x_1$ as the only predictor and plotting number of games won $y$ and
    versus the team's rushing yards $x_1$, we see that $x_1$ is a significant predictor and is linearly related to the number of wins.
    
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    linear_model <- lm(y ~ x1, data = data_set)
    summarize_linear_model(linear_model)
    ggplot(data_set, aes(x = x1, y = y)) +
        geom_point(alpha = 0.2) +
        geom_smooth(method = "lm", se = FALSE) +
        labs(
            x = "rushing yards",
            y = "number of games won",
            title = "Number of Games Won vs. Rushing Yards"
        ) +
        theme(
            plot.title = element_text(hjust = 0.5),
            axis.text.x = element_text(angle = 0)
        )
    ```
    
    $x_1$ is linearly related to the response on its own. That being said, it is not needed in a model with $x_2$, $x_7$, and $x_8$
    as it doesn't improve the predictions significantly as it is moderately correlated with $x_8$ and highly correlated with $x_7$ /
    doesn't provide much addition insight to the prediction when the other predictors are already in the model.
    
    ```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
    library(dplyr)
    cor(data_set %>% select(y, x1, x2, x7, x8))
    ```