---
title: "hw11"
author: "Tom Lever"
date: 11/11/22
output:
  pdf_document: default
  html_document: default
urlcolor: blue
linkcolor: red
---

# Metadata

Name:    hw11

URL:     https://github.com/tslever/DS5100-2022-08-tsl2b/blob/main/lessons/M11/hw11.Rmd

Course:  DS 5100

Term:    Fall 2022 Online

Module:  M11: R Programming

Topic:   Tidyverse

Author:  Tom Lever

Net ID:  tsl2b

Created: 11 November 2022

Updated: 11 November 2022

# Overview

In this homework, you will work on the Abalone datset from the UCI Machine Learning Repository.

To get started, download and import the `abalone.data` dataset from this URL:

* [https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data]( https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data)

You can pass the URL directly to `read.csv`. There is no header row.

Note: The instruction to print in the questions below can be accomplished either through the `print` function or by displaying a value directly.

**Total Points: 7**

# Tasks

## Task 0
(0 points)

Get the dataset.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
data_set <- read.csv(
   'https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data',
   header = FALSE
)
head(data_set, n = 3)
```

## Task 1
(1 point)

Print the number of rows in the dataset.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
print(nrow(data_set))
```

## Task 2
(1 point)

The rightmost column is the number of rings. Print the maximum number of rings.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
library(dplyr)
print(data_set %>% select(V9) %>% max())
```

## Task 3
(1 point)

The leftmost column is the gender with values `M` (Male), `F` (Female), and `I` (Infant).

Apply the `filter` function from `tidyverse` to select only rows where gender is infant, and print the number of records.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
print(data_set %>% filter(V1 == 'I') %>% nrow())
```

## Task 4
(1 point)

Apply the `filter` function from `tidyverse` to select only rows where gender is infant or male, and print the number of records.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
print(data_set %>% filter (V1 == 'I' | V1 == 'M') %>% nrow())
```

## Task 5
(1 point)

Call the `table` function on the abalone genders to find out how many of each gender are present.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
data_set %>% select(V1) %>% table()
```

## Task 6
(1 point)

Compute the mean value of column 2 (V2) grouped by gender.

V2 is the longest shell measurement.

Requirements: Use the `%>%` operator to chain commands, and the `group_by` and `summarize` functions.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
data_set %>% group_by(V1) %>% summarize(average = mean(V2))
```

## Task 7
(1 point)

Compute the median value of longest shell measurement for only the males.

Requirements: Use the `%>%` operator to chain commands.

```{r, eval=TRUE, echo=TRUE, results="show", message=FALSE}
data_set %>% filter(V1 == 'M') %>% summarize(the_median = median(V2))
data_set %>% group_by(V1) %>% summarize(the_median = median(V2)) %>% filter(V1 == 'M')
```